---
title: "Descriptives and Data Exploration - International Aid"
author: "Saewon Park"
date: "June 30, 2020"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse,
               corrplot,
               shiny,
               naniar,
               viridis)


```


## Purpose

This analysis explores the dataset created by "aid_data_cleaning.Rmd" in order to understand the data and plan for an appropriate shiny dashboard to present initial insights from the data in the most efficient way.



### Loading the data

```{r data import}

#importing the final dataset created by "aid_data_cleaning.Rmd"
merged_dyadic <- read.csv("./processed_data/merged_dyadic_analysis_data.csv",
                          stringsAsFactors = F) 


#Note: first I used lagged country-level controls and then changed to unlagged so that all variables (country-level or dyad-level) match on year

```


## Descriptives & Data Exploration

```{r summary stats}

#to show more data in the console
options(max.print=4000)

#just seeing the summary of all variables
summary(merged_dyadic)
  #note that all country level controls are already lagged one year (think about whether to change this or not)
```




```{r correlation}


#create a df without the lags easier exploratory analysis 
  #so that map functions for all numeric variables can be used easily 
  #remove lags, logs, and other transformations
numeric_dyadic <- merged_dyadic %>%
  ungroup() %>% #ungrouping because I won't necessarily use ccodes for plotting
  select_at(c(vars(-matches("(_(\\d)+|_lag|_log|_squared|_corrected|_fd|_sum)$")),
              vars(-matches("^(commit_|distw)")))) %>%
  mutate(dyad_name = paste0(d_country, " - ", r_country)) #create human-readable dyad names

#corrplot
corr_all_merged <- cor(select(numeric_dyadic, #cor object to feed into corrplot
                              -c(ccode1,
                                 ccode2,
                                 r_country,
                                 d_country,
                                 r_region,
                                 d_region,
                                 dyad_id,
                                 dyad_name)), 
                       use = "pairwise.complete.obs")


corrplot(corr_all_merged,
         type = "upper") #no surprises here


```



```{r trendlines}

#trendlines (dyads as obs, year as x var, each var as y var)



#transforming the df so that I can facet by the numeric variables
merged_dyadic_plot <- numeric_dyadic %>%
  select(-c(ccode1, #remove non-numeric vars except for dyad_name, (d_/r_)region and (d_/r_)country
            ccode2,
            dyad_id,
            deflator)) %>%
  pivot_longer(-c(dyad_name,
                  year,
                  r_region,
                  d_region,
                  r_country,
                  d_country,
                  contig,
                  comlang_off,
                  comlang_ethno,
                  colony,
                  dist),
               names_to = "numeric_var",
               values_to = "value")
  

#checking that the faceted plot works by using a sample (because the whole data is very big)
merged_dyadic_plot_sample <- merged_dyadic_plot %>%
  filter(d_region == "Europe & Central Asia")
  
  
trendline_exploratory_sample_gplot <- ggplot(merged_dyadic_plot_sample) +
  geom_line(aes(x = year,
                y = value,
                color = dyad_name)) +
  facet_wrap(~ numeric_var,
             scales = "free_y") +
  theme_bw() + 
  theme(legend.position = "none") 

trendline_exploratory_sample_gplot


#I think some variables need to be points or other visualizations
#I think donor and recipient and dyadic variables should be in separate sections (for donor and recipient they can be compared next to each other)

#I think it may be a good idea to show some trendlines as a mini dashboard 
  # Goal: See and explore the trends and relationship between economic factors and aid by regions.
```


```{r missing data}

#not using Amelia because it is heavy and crashes R. Plus, we're not doing imputation at this point

#other packages:
  #naniar by a stats PhD: https://cran.r-project.org/web/packages/naniar/vignettes/naniar-visualisation.html
    #the plots are prettier
  #VIM as recommended by Datacamp: https://www.datacamp.com/community/tutorials/visualize-data-vim-package
    # the parallel coordinate plot (parcoordMiss()) is pretty cool

#let's try to use naniar first
  #the data is quite big, so I divide into recipient variables, donor variables, and dyadic variables


#recipient variables
gg_miss_fct(numeric_dyadic %>% 
           filter(d_region == "North America") %>% #reducing the data so that the function works
           select_at(c(vars(matches("^r_")),
                       vars(year))),
           year) +  
  scale_fill_viridis(option = "D", #overriding naniar's fill setting to choose a diff gradient
                     direction = -1,
                     end = 0.75) +
  labs(title = "Missing data in recipient-side variables",
       fill = "% of data missing")

#donor variables
gg_miss_fct(numeric_dyadic %>% 
           filter(d_region == "North America") %>% #reducing the data so that the function works
           select_at(c(vars(matches("^d_")),
                       vars(year))),
           year) + 
  scale_fill_viridis(option = "D", 
                     direction = -1,
                     end = 0.75) +
  labs(title = "Missing data in donor-side variables",
       fill = "% of data missing")


#dyadic variables
gg_miss_fct(numeric_dyadic %>% 
           filter(d_region == "North America") %>% #reducing the data so that the function works
           select_at(c(vars(-matches("^(r_|d_)")),
                       vars(year))),
           year) + 
  scale_fill_viridis(option = "D", 
                     direction = -1,
                     end = 0.75) +
  labs(title = "Missing data in dyadic variables",
       fill = "% of data missing")

  



#Insights:
  # around 1990 is the cutoff where is a significant change in the level of missing data
  # unemployment has 0 data before 1990s -> we may need a better source
  # recipient countries have substantially more missing data 

#I can make a little blog about this data inequality 

```




```{r econ mini dash ggplots}

#first mini dashboard for trendlines

# donor vars: d_pop, d_rgdpe_per_capita, d_unemployment
# recipient: r_pop, r_rdgde_per_capita, r_unemployment
  # remember that (r_/d_)pop are not logged and in millions
# aid vars: disburse_*_aid
  # be able choose to visually group by donor or recipient region
# customizable variables: year range, donor region, recipient region
# nice to have
    # important dates on timeline (ex- 9/11)
    # show country/dyad name on hover (maybe a must have)


#due to missing data (especially for unemployment) before 1990s, let's set the year range to 1991
  #if unemployment and trade missing data are fixed (by funding different datasets, we can increase the range)


#donor vars
econ_dash_donor <- merged_dyadic_plot %>%
  select(year,
         d_country,
         d_region,
         numeric_var,
         value) %>%
  filter(numeric_var %in% c("d_pop",
                            "d_rgdpe_per_capita",
                            "d_unemployment") &
           !is.na(d_region) &
           year %in% c(1991:2015)) %>% #years where most variable have data (also cuts data size by around 40%)
  mutate(numeric_var = case_when(numeric_var == "d_pop" ~ "Population (in millions)",
                                 numeric_var == "d_rgdpe_per_capita" ~ "GDP per capita (in constant USD)",
                                 numeric_var == "d_unemployment" ~ "Unemployment (%)"),
         d_region = factor(d_region,
                           levels = c("East Asia & Pacific",
                                      "Europe & Central Asia",
                                      "Latin America & Caribbean",
                                      "Middle East & North Africa",
                                      "South Asia",
                                      "Sub-Saharan Africa",
                                      "North America"))) %>%
  distinct() %>%
  group_by(year,
           d_region,
           numeric_var) %>%
  summarise(value = sum(value,
                        na.rm = T)) 

econ_dash_donor_gplot <- ggplot(econ_dash_donor) +
  geom_line(aes(x = year,
                y = value,
                color = d_country)) +
  facet_wrap(~ numeric_var,
             scales = "free_y") +
  theme_bw() + 
  theme(legend.position = "none") 

#econ_dash_donor_gplot


#recipient vars
econ_dash_recipient <- merged_dyadic_plot %>%
  select(year,
         r_country,
         r_region,
         numeric_var,
         value) %>%
  filter(numeric_var %in% c("r_pop",
                            "r_rgdpe_per_capita",
                            "r_unemployment") &
           !is.na(r_region) &
           year %in% c(1991:2015)) %>% #years where most variable have data (also cuts data size by around 40%)
  mutate(numeric_var = case_when(numeric_var == "r_pop" ~ "Population (in millions)",
                                 numeric_var == "r_rgdpe_per_capita" ~ "GDP per capita (in constant USD)",
                                 numeric_var == "r_unemployment" ~ "Unemployment (%)"),
         r_region = factor(r_region,
                           levels = c("East Asia & Pacific",
                                      "Europe & Central Asia",
                                      "Latin America & Caribbean",
                                      "Middle East & North Africa",
                                      "South Asia",
                                      "Sub-Saharan Africa",
                                      "North America"))) %>%
  distinct() %>%
  group_by(year,
           r_region,
           numeric_var) %>%
  summarise(value = sum(value,
                        na.rm = T)) 

econ_dash_recipient_gplot <- ggplot(econ_dash_recipient) +
  geom_line(aes(x = year,
                y = value,
                group = r_country,
                color = r_region)) +
  facet_wrap(~ numeric_var,
             scales = "free_y") +
  theme_bw() + 
  theme(legend.position = "none") 

#econ_dash_recipient_gplot


#dyadic vars
econ_dash_dyadic <- merged_dyadic_plot %>%
  filter(grepl("disburse_.+_aid", 
               numeric_var) &
           numeric_var != "disburse_other_aid" &
           year %in% c(1991:2015)) %>% #years where most variable have data (also cuts data size by around 40%)
  mutate(numeric_var = case_when(numeric_var == "disburse_democracy_aid" ~ "Democracy Aid",
                                 numeric_var == "disburse_econ_capacity_aid" ~ "Economic Capacity Aid",
                                 numeric_var == "disburse_social_aid" ~ "Social Aid",
                                 numeric_var == "disburse_disaster_aid" ~ "Disaster Aid",
                                 numeric_var == "disburse_production_aid" ~ "Production Aid",
                                 numeric_var == "disburse_program_aid" ~ "Program Aid")) 

econ_dash_dyadic_gplot <- ggplot(econ_dash_dyadic) +
  geom_line(aes(x = year,
                y = value,
                group = dyad_name, #group splits the data like color or fill but does not apply additional visual differentiation
                color = r_region)) + 
  facet_wrap(~ numeric_var,
             scales = "free_y") +
  theme_bw() 

#econ_dash_dyadic_gplot


#dyadic vars - summarized into regions (to make it lighter)

econ_dash_dyadic_region <- econ_dash_dyadic %>% 
  select(year,
         r_region,
         d_region,
         numeric_var,
         value) %>%
  filter(!(is.na(r_region) | is.na(d_region))) %>%
  mutate(region_dyad = paste0(d_region, " to ", r_region),
         r_region = factor(r_region,
                           levels = c("East Asia & Pacific",
                                      "Europe & Central Asia",
                                      "Latin America & Caribbean",
                                      "Middle East & North Africa",
                                      "South Asia",
                                      "Sub-Saharan Africa",
                                      "North America")),
         d_region = factor(d_region,
                           levels = c("East Asia & Pacific",
                                      "Europe & Central Asia",
                                      "Latin America & Caribbean",
                                      "Middle East & North Africa",
                                      "South Asia",
                                      "Sub-Saharan Africa",
                                      "North America"))) %>%
  group_by(year,
           r_region,
           d_region,
           region_dyad,
           numeric_var) %>%
  summarise(value = sum(value)) 

econ_dash_dyadic_region_gplot <- ggplot(econ_dash_dyadic_region) +
  geom_line(aes(x = year,
                y = value,
                group = region_dyad, #group splits the data like color or fill but does not apply additional visual differentiation
                color = r_region)) + 
  facet_wrap(~ numeric_var,
             scales = "free_y") +
  theme_bw() 

#econ_dash_dyadic_region_gplot


#setting my own colors for each region (using Set1 colors)
region_cols <- c(`East Asia & Pacific` = "#E41A1C",
                 `Europe & Central Asia` = "#377EB8",
                 `Latin America & Caribbean` = "#4DAF4A",
                 `Middle East & North Africa` = "#984EA3",
                 `South Asia` = "#FF7F00",
                 `Sub-Saharan Africa` = "#E7298A",
                 `North America` = "#A65628")



```


```{r shiny functions}

#new functions to limit min/max date views
  #source: https://stackoverflow.com/questions/31152960/display-only-months-in-daterangeinput-or-dateinput-for-a-shiny-app-r-programmin

dateInput2 <- function(inputId, label, minview = "days", maxview = "decades", ...) {
  d <- shiny::dateInput(inputId, label, ...)
  d$children[[2L]]$attribs[["data-date-min-view-mode"]] <- minview
  d$children[[2L]]$attribs[["data-date-max-view-mode"]] <- maxview
  d
}

dateRangeInput2 <- function(inputId, label, minview = "days", maxview = "decades", ...) {
  d <- shiny::dateRangeInput(inputId, label, ...)
  d$children[[2L]]$children[[1]]$attribs[["data-date-min-view-mode"]] <- minview
  d$children[[2L]]$children[[3]]$attribs[["data-date-min-view-mode"]] <- minview
  d$children[[2L]]$children[[1]]$attribs[["data-date-max-view-mode"]] <- maxview
  d$children[[2L]]$children[[3]]$attribs[["data-date-max-view-mode"]] <- maxview
  d
}


```


####first mini dashboard for trendlines

- donor vars: d_pop, d_rgdpe_per_capita, d_unemployment
- recipient: r_pop, r_rdgde_per_capita, r_unemployment
  - remember that (r_/d_)pop are not logged and in millions
- aid vars: disburse_*_aid
  - be able choose to visually group by donor or recipient region
- customizable variables: year range, donor region, recipient region
- nice to have
    - important dates on timeline (ex- 9/11)
    - show country/dyad name on hover (maybe a must have)
-due to missing data (especially for unemployment) before 1990s, let's set the year range to 1991
  -if unemployment and trade missing data are fixed (by funding different datasets, we can increase the range)

-reference: 
  -https://shiny.rstudio.com/tutorial/written-tutorial/lesson1/
  -https://shiny.rstudio.com/articles/layout-guide.html
  -shiny's gallery for each function is helpful


```{r econ mini dash}


#define UI for an app that draws economic and aid trend graphs
ui <- fluidPage(
  
  #app title 
  titlePanel("Economic Trends and Aid Distribution"),
  
  fluidRow(
    column(3,
           h4("Dashboard Parameters"),
           
           #input: date range for the years to be shown
           dateRangeInput2(inputId = "yearRange",
                          label = "Year range:",
                          start = "1991-01-01",
                          end = "2015-12-31",
                          min = "1991-01-01",
                          max = "2015-12-31",
                          format = "yyyy",
                          startview = "decade",
                          minview = "decade"
                          ),
           
           selectInput(inputId = "recipientRegion",
                       label = "Recipient country region:",
                       choices = c("All regions",
                                   na.omit(unique(econ_dash_recipient$r_region))),
                       selected = "All regions"),
           
           selectInput(inputId = "donorRegion",
                       label = "Donor country region:",
                       choices = c("All regions",
                                   na.omit(unique(econ_dash_donor$d_region))),
                       selected = "All regions")
           ),
    column(9,
           h3("International Aid Trends by Type"),
           
           radioButtons(inputId = "aidColor",
                        label = "Label/color by:",
                        choices = list("Recipient region" = "r_region",
                                       "Donor region" = "d_region")),
           
           plotOutput(outputId = "aidTrends")
           )
    ),
  
  h3("Economic Trends"),
  plotOutput(outputId = "recipientTrends",
             height = "200px"),
  plotOutput(outputId = "donorTrends",
             height = "200px")
)
  

server = function(input, output) {
  
  #creating reactive variables to change x axis limits (years)
  yearStart <- reactive({
    input$yearRange[1] %>%
      format(.,
             "%Y") %>%
      as.numeric(.)
    })
    
  yearEnd <- reactive({
    input$yearRange[2] %>%
      format(.,
             "%Y") %>%
      as.numeric(.)
  })
    

  #renderPlot is used to indicate that: 
    #1. the output is of the plot type
    #2. it is "reactive" and therefore should be automatically re-executed when inputs change
  output$aidTrends <- renderPlot({ #this takes forever to load
    
    if (input$recipientRegion != "All regions") { #applying region filters
      econ_dash_dyadic_region <- econ_dash_dyadic_region %>%
        filter(r_region == input$recipientRegion)
    }
    
    if (input$donorRegion != "All regions") {
      econ_dash_dyadic_region <- econ_dash_dyadic_region %>%
        filter(d_region == input$donorRegion)
    } 
    
    if (input$aidColor == "r_region") {
    aidTrendsGG <- ggplot(econ_dash_dyadic_region) +
      geom_line(aes(x = year,
                    y = value,
                    group = region_dyad, #group splits the data like color or fill but does not apply additional visual differentiation
                    color = r_region)) 
    } else {
    aidTrendsGG <- ggplot(econ_dash_dyadic_region) +
      geom_line(aes(x = year,
                    y = value,
                    group = region_dyad, #group splits the data like color or fill but does not apply additional visual differentiation
                    color = d_region))
    }
    
    aidTrendsGG + 
      xlim(yearStart(),
           yearEnd()) +
      facet_wrap(~ numeric_var,
                 scales = "free_y") +
      labs(color = "Regions") +
      scale_color_manual(values = region_cols,
                         drop = F) + 
      theme_bw() 
      
  })
  
  output$recipientTrends <- renderPlot({
    
    if (input$recipientRegion != "All regions") { #applying region filters
      econ_dash_recipient <- econ_dash_recipient %>%
        filter(r_region == input$recipientRegion)
    }
    
    recipientTrendsGG <- ggplot(econ_dash_recipient) +
      geom_line(aes(x = year,
                    y = value,
                    group = r_region,
                    color = r_region)) +
      xlim(yearStart(),
           yearEnd()) +
      facet_wrap(~ numeric_var,
                 scales = "free_y") +
      scale_color_manual(values = region_cols) +
      theme_bw() + 
      theme(legend.position = "none") 
    recipientTrendsGG
  }) 
  
  output$donorTrends <- renderPlot({
    
    if (input$donorRegion != "All regions") { #applying region filters
      econ_dash_donor <- econ_dash_donor %>%
        filter(d_region == input$donorRegion)
    } 
    
    donorTrendsGG <- ggplot(econ_dash_donor) +
      geom_line(aes(x = year,
                    y = value,
                    group = d_region,
                    color = d_region)) +
      xlim(yearStart(),
           yearEnd()) +
      facet_wrap(~ numeric_var,
                 scales = "free_y") +
      scale_color_manual(values = region_cols) +
      theme_bw() + 
      theme(legend.position = "none") 
    donorTrendsGG
  })

}

shinyApp(ui = ui, server = server)




```

Hosting Shiny on sites:

  1. https://shiny.rstudio.com/tutorial/written-tutorial/lesson7/#:~:text=The%20easiest%20way%20to%20turn,app%20including%20server%20administration%20tools.

  2. (Use Google account) https://www.shinyapps.io/admin/#/dashboard 


## Plots for map dashboard

```{r ggplots}

#first build the individual plots that will go in the dashboard and then combine them into a dashboard
  #1. Map plot (doesn't have to be an exact map, but needs to visually show all countries, and would be useful if they are somewhat in their geographical positions)
    #some resources:
      #https://www.r-graph-gallery.com/hexbin-map.html
      #https://www.gapminder.org/tools/ 
      #https://flowingdata.com/
      #could be a circular visualization or a variation of it (circle with network visualization)
  #2. Top 5 plot/list
  #3. Overall volume trend plot




```

