---
title: "What determines the flow of democracy aid between donor states and recipient states?"
author: "SaewonPark"
date: "December 6, 2018"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

pacman::p_load(stargazer)

```

## Background

- Aid is often considered in the aggregate and the issues of this have been explored
- Still, many studies on the factors of aid-link formation still see aid as a reward and build mechanisms on that.
  - even studies that differentiate between the types of aid.
  
## Democracy Aid

- Definition

## Previous Reseach

- Aid (including democracy aid) as a reward for
  - Foreign policy alignment
  - Economic ties
  - Altruistic/normative reasons
  
## Argument

- I argue that the reward explanation does not apply to democracy aid because:
  1. Condition to a reward rather than the reward
  2. Can pose a threat to the regime (direct challenge to regime or sovereignty)
  3. Fewer fungible components

## Argument

- Then, why do bilateral democracy aid links form?
  - This is where I am still working
  
  - Immigration (normative) *
    - diaspora can be a signal from donor side, but it is also linked to recipient signal because of the link between the diaspora and their families/friends
    - the strength of ties of diaspora to country of origin as a predictor? (need to isolate effect to democracy aid only)

  - Military aid (few important cases)
  
  - As a signal to receive other aid (unobservable and endogeneous)
  - Signals from the recipient side in general that signifies a want for democracy and a window of opportunity. 


## Research Design


## Control

  - Other types of aid
  
  - Structural effects of existing aid network (social network effects - but not a causal argument)
    - donorXdonor or recipientXrecipient matrices of shared donors/recipients, eigenvector centrality

  - Level of democracy
  
## Previous result that focused on UNGA voting patterns 

- this was more methodological than causal


## Data

```{r, echo = F, results= 'asis'}

mergepdata <- read.csv("UNGA_fulldata.csv", stringsAsFactors = F)

#create a dataframe with only the variables that will be shown in the descriptives
mergepdata_desc <- dplyr::select(mergepdata, 
                                 aidflow_corrected,
                                 s3un, 
                                 ipfd, 
                                 r_conflict_prioryear,
                                 r_politychange,
                                 democratization,
                                 r_democracy,
                                 r_gdp_growth,
                                 d_gdp_growth,
                                 r_gdplog, 
                                 d_gdplog,
                                 r_population,
                                 d_population,
                                 d_women_parliament,
                                 comlang_off,
                                 colony,
                                 distlog,
                                 ptacount_1,
                                 intervention_1,
                                 r_unemployment_1,
                                 d_unemployment_1,
                                 unempdiff_1,
                                 logimmigration_1,
                                 logrtod_export_1,
                                 logdtor_export_1)



stargazer(mergepdata_desc,
          type = "latex",
          header = F,
          covariate.labels = c("Aidflow (per 1000, logged)",
                               "Ideal Point Difference",
                               "Change in Ideal Point Difference",
                               "Conflict, Prior Year, R",
                               "Democratization (ord), R",
                               "Democratization (dummy), R",
                               "Democracy (ord), R",
                               "GDP growth, R",
                               "GDP growth, D",
                               "GDP (logged), R",
                               "GDP (logged), D",
                               "Population (logged), R",
                               "Population (logged), D",
                               "Women in Parliament (Percent), D",
                               "Common Official Lang.",
                               "Past Colonial Relationship",
                               "Distance (km, logged)",
                               "# of PTAs",
                               "Military Intervention (dummy)",
                               "Unemployment Rate, R",
                               "Unemployment Rate, D",
                               "Unemployment Difference",
                               "Immigration (logged), R to D",
                               "Export (logged), R to D",
                               "Export (logged), D to R"),
          title = "Descriptive Statistics, 1991 - 2014")


```




## Questions

- Immigration is correlated with other types of aid as well
  - How can I isolate the effects of transnational democracy movements led by diaspora?
  
- Should I focus on just either military aid or immigration?



- Research question and contribution to the literature
- Argument
- Research Design
- Estimation
- Results
- Open questions/issues/problems




