---
title: "ma_essay_models"
author: "SaewonPark"
date: "December 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


pacman::p_load(tidyverse,
               countrycode,
               haven,
               readxl,
               plm,
               corrplot,
               lmtest,
               multiwayvcov,
               car,
               stargazer,
               jtools,
               dynsim)

lag <- dplyr::lag

```


Set-up for regression models

Dec 29 - limited data to 1995 - 2014
due to coverage
```{r}

#pdata frame
mergepdata4 <- pdata.frame(select(mergedata4,
                                  -c(disaster_aid_com_corr_1,
                                     econ_capacity_aid_com_corr_1,
                                     other_aid_com_corr_1,
                                     production_aid_com_corr_1,
                                     program_aid_com_corr_1,
                                     social_aid_com_corr_1,
                                     democracy_aid_disb_corr_1,
                                     disaster_aid_disb_corr_1,
                                     econ_capacity_aid_disb_corr_1,
                                     other_aid_disb_corr_1,
                                     production_aid_disb_corr_1,
                                     program_aid_disb_corr_1,
                                     social_aid_disb_corr_1,
                                     disaster_aid_com_corr_1_lag,
                                     econ_capacity_aid_com_corr_1_lag,
                                     other_aid_com_corr_1_lag,
                                     production_aid_com_corr_1_lag,
                                     program_aid_com_corr_1_lag,
                                     social_aid_com_corr_1_lag,
                                     democracy_aid_disb_corr_1_lag,
                                     disaster_aid_disb_corr_1_lag,
                                     econ_capacity_aid_disb_corr_1_lag,
                                     other_aid_disb_corr_1_lag,
                                     production_aid_disb_corr_1_lag,
                                     program_aid_disb_corr_1_lag,
                                     social_aid_disb_corr_1_lag,
                                     r_pop_notlogged_1,
                                     r_politychange_1,
                                     r_democracy_1,
                                     r_gdp_growth_1,
                                     r_gdp_1,
                                     r_population_1,
                                     r_democratization_1,
                                     r_gdplog_1,
                                     r_democracy_sq_1,
                                     d_pop_notlogged_1,
                                     d_gdp_1,
                                     d_population_1,
                                     d_consumer_inflation_1,
                                     d_women_parliament_1,
                                     d_gdplog_1,
                                     r_pop_notlogged_1_lag,
                                     r_politychange_1_lag,
                                     r_democracy_1_lag,
                                     r_gdp_growth_1_lag,
                                     r_gdp_1_lag,
                                     r_population_1_lag,
                                     r_democratization_1_lag,
                                     r_gdplog_1_lag,
                                     r_democracy_sq_1_lag,
                                     d_pop_notlogged_1_lag,
                                     d_gdp_1_lag,
                                     d_population_1_lag,
                                     d_consumer_inflation_1_lag,
                                     d_women_parliament_1_lag,
                                     d_gdplog_1_lag,
                                     ptasum_1,
                                     ptacount_1,
                                     r_unemployment_1,
                                     d_unemployment_1,
                                     unempdiff_1,
                                     immigration_1,
                                     naturalization_1,
                                     rtod_export_1,
                                     dtor_export_1,
                                     log_rtod_export_1,
                                     log_dtor_export_1,
                                     military_aid_corr_1,
                                     ptasum_1_lag,
                                     ptacount_1_lag,
                                     r_unemployment_1_lag,
                                     d_unemployment_1_lag,
                                     unempdiff_1_lag,
                                     immigration_1_lag,
                                     naturalization_1_lag,
                                     rtod_export_1_lag,
                                     dtor_export_1_lag,
                                     log_rtod_export_1_lag,
                                     log_dtor_export_1_lag,
                                     military_aid_corr_1_lag)) %>%
                             filter(year > 1994),
           index = c("ccode1", "ccode2"))
#this has all the variables
#dec 29 - removed some duplicate variables
```


correlation
```{r}
corrvar <- select(mergepdata4 %>%
                  ungroup(),
                  democracy_aid_com_corr,
                  ipdiff, 
                  ipfd, 
                  r_conflict_prioryear,
                  r_politychange,
                  r_democracy,
                  #r_democratization,
                  r_gdp_growth,
                  d_gdp_growth,
                  r_gdplog, 
                  d_gdplog,
                  r_population,
                  d_population,
                  d_women_parliament,
                  comlang_off,
                  colony,
                  distlog,
                  ptacount,
                  ptasum,
                  r_unemployment,
                  d_unemployment,
                  unempdiff,
                  log_immigration,
                  log_naturalization,
                  log_rtod_export,
                  log_dtor_export,
                  other_aid_com_corr,
                  military_aid_corr)

corr_tbl <- cor(corrvar, use = "complete.obs")

# png(filename="corrplot.png",
#     units="in", 
#     width=9, 
#     height=9, 
#     res=144)
corrplot(corr_tbl, method = "number")
# dev.off()


# immigration is correlated with naturalization: 0.71
  # it is even lower in the full sample of just immigration data (0.5)
  # if I regress two variables that are correlated, I will get a significant result but without a theory, it doesn't matter as much

# I also look at VIF in the bottom
```






```{r}
#descriptive stats
stargazer(as.data.frame(corrvar),
          type = "latex",
          header = F,
          covariate.labels = c("Democracy Aidflow (per 1000, logged)",
                               "Ideal Point Difference",
                               "Change in Ideal Point Difference",
                               "Conflict, Prior Year, R",
                               "Democratization (ord), R",
                               #"Democratization (dummy), R",
                               "Democracy (ord), R",
                               "GDP growth, R",
                               "GDP growth, D",
                               "GDP (logged), R",
                               "GDP (logged), D",
                               "Population (logged), R",
                               "Population (logged), D",
                               "Women in Parliament (Percent), D",
                               "Common Official Lang.",
                               "Past Colonial Relationship",
                               "Distance (km, logged)",
                               "# of PTAs",
                               "Cumulative Sum PTAs",
                               #"Military Intervention (dummy)",
                               "Unemployment Rate, R",
                               "Unemployment Rate, D",
                               "Unemployment Difference",
                               "Immigration (logged), R to D",
                               "Naturalization (logged), R to D",
                               "Export (logged), R to D",
                               "Export (logged), D to R",
                               "US Military Aid (per 1000, logged)",
                               "Other Economic Aid (per 1000, logged)"),
          title = "Descriptive Statistics, 1995 - 2014")



```


**Dec 29
I think maybe I should remove GDP growth and make democratization binary


Base Model
```{r}
#all covariates lagged 1

base_m <- lm(democracy_aid_com_corr ~ 
               log_naturalization_lag +
                ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag +
                r_democracy_sq_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                #log_naturalization_roll_lag +
                #naturalization_prop_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4,
              na.action = na.omit)

#summary(base_m)

options(scipen = 999)
base_m_hc3 <- coeftest(base_m, cluster.vcov(base_m,
                                            cluster = cbind(mergepdata4$ccode1, mergepdata4$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(22:nrow(summary(base_m)$coefficients)),]; base_m_hc3

#naturalization survives despite hc3-two way clustered errors
#democracy (+sq) and democratization were significant and in the correct direction *without* clustered errors

#even with immigration included in the model, naturalization survives, although the vif for immigration and naturalization are high (7-8)

#rolling sum (5 year) of naturalization proportional to population is in the right direction but it is not significant (p is around 0.3)

  # however, it is very significant and holds the right direction when it is not proportional to population
  # when proportinal to pop, it is very small and not logged


```



```{r}
#base model with other aid types 

base_m_aid <- lm(democracy_aid_com_corr ~ 
                   log_naturalization_lag +
                   ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag +
                r_democracy_sq_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                #log_immigration_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                other_economic_aid_lag +
                military_aid_corr_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4,
              na.action = na.omit)

#summary(base_m_aid)

#all aid except for military aid affects democracy aid (positively)

base_m_aid_hc3 <- coeftest(base_m_aid, cluster.vcov(base_m_aid,
                                            cluster = cbind(mergepdata4$ccode1, mergepdata4$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(24:nrow(summary(base_m_aid)$coefficients)),]; base_m_aid_hc3

#naturalization survives with hc3 and two way clusters

```

pre 2001 
```{r}

mergepdata4_pre <- mergepdata4 %>%
  filter(year < 2002)

#base model with other aid types 

base_m_aid_pre <- lm(democracy_aid_com_corr ~ 
                   log_naturalization_lag +
                   ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag +
                r_democracy_sq_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                #log_immigration_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                other_economic_aid_lag +
                military_aid_corr_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4_pre,
              na.action = na.omit)

#summary(base_m_aid)

#all aid except for military aid affects democracy aid (positively)

base_m_aid_pre_hc3 <- coeftest(base_m_aid_pre, cluster.vcov(base_m_aid_pre,
                                            cluster = cbind(mergepdata4_pre$ccode1, mergepdata4_pre$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(24:nrow(summary(base_m_aid_pre)$coefficients)),]; base_m_aid_pre_hc3

#naturalization survives with hc3 and two way clusters before 2001

```

post 2001
```{r}

mergepdata4_post <- mergepdata4 %>%
  filter(year > 2001)

#base model with other aid types 

base_m_aid_post <- lm(democracy_aid_com_corr ~ 
                   log_naturalization_lag +
                   ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag +
                r_democracy_sq_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                #log_immigration_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                other_economic_aid_lag +
                military_aid_corr_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4_post,
              na.action = na.omit)

#summary(base_m_aid)

#all aid except for military aid affects democracy aid (positively)

base_m_aid_post_hc3 <- coeftest(base_m_aid_post, cluster.vcov(base_m_aid_post,
                                            cluster = cbind(mergepdata4_post$ccode1, mergepdata4_post$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(24:nrow(summary(base_m_aid_post)$coefficients)),]; base_m_aid_post_hc3

#naturalization survives with hc3 and two way clusters before 2001

```


without irq (645) or afgh (700)

```{r}

mergepdata4_out <- mergepdata4 %>%
  filter(!(ccode2 %in% c(645, 700)))

#base model with other aid types 

base_m_aid_out <- lm(democracy_aid_com_corr ~ 
                   log_naturalization_lag +
                   ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag +
                r_democracy_sq_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                #log_immigration_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                other_economic_aid_lag +
                military_aid_corr_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4_out,
              na.action = na.omit)

#summary(base_m_aid)

#all aid except for military aid affects democracy aid (positively)

base_m_aid_out_hc3 <- coeftest(base_m_aid_out, cluster.vcov(base_m_aid_out,
                                            cluster = cbind(mergepdata4_out$ccode1, mergepdata4_out$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(24:nrow(summary(base_m_aid_out)$coefficients)),]; base_m_aid_out_hc3

#naturalization survives with hc3 and two way clusters after 2001

```


without outliers that were found in influence plot
France (220), Netherlands (210), Austria (305), Japan (740)
Qatar (694), Bhutan (760), Cape Verde (402), China (710)

```{r}

mergepdata4_out2 <- mergepdata4 %>%
  filter(!(ccode1 %in% c(220, 210, 305, 740) | ccode2 %in% c(694, 760, 402, 710)))

#base model with other aid types 

base_m_aid_out2 <- lm(democracy_aid_com_corr ~ 
                   log_naturalization_lag +
                   ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag +
                r_democracy_sq_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                #log_immigration_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                other_economic_aid_lag +
                military_aid_corr_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4_out2,
              na.action = na.omit)

#summary(base_m_aid)

#all aid except for military aid affects democracy aid (positively)

base_m_aid_out2_hc3 <- coeftest(base_m_aid_out2, cluster.vcov(base_m_aid_out2,
                                            cluster = cbind(mergepdata4_out2$ccode1, mergepdata4_out2$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(24:nrow(summary(base_m_aid_out2)$coefficients)),]; base_m_aid_out2_hc3

#naturalization survives with hc3 and two way clusters when outliers are removed

# I think it's better to just include the first outlier model since I don't really have a way to explain this

```



Latex tables
```{r}

diagnostics <- c("R-squared", "F-stat", "Obs.")

fit.base <- summary(base_m, diagnostics = T)
fit.base_aid <- summary(base_m_aid, diagnostics = T)
fit.pre <- summary(base_m_aid_pre, diagnostics = T)
fit.post <- summary(base_m_aid_post, diagnostics = T)
fit.out <- summary(base_m_aid_out, diagnostics = T)


stargazer(base_m_hc3,
          base_m_aid_hc3,
          base_m_aid_pre_hc3,
          base_m_aid_post_hc3,
          base_m_aid_out_hc3,
          type = "latex",
           covariate.labels = c("Naturalization (logged), R to D",
                                "Ideal Point Diff. (5 year lag)",
                               "Conflict, Prior Year, R",
                               "Democratization (ord), R",
                               "Democracy (ord), R",
                               "Democracy (squared, ord), R",
                               "GDP growth, R",
                               "GDP growth, D",
                               "GDP (logged), R",
                               "GDP (logged), D",
                               "Population (logged), R",
                               "Population (logged), D",
                               "Women in Parliament (%), D",
                               "Common Official Lang.",
                               "Past Colonial Relationship",
                               "Distance (km, logged)",
                               "# of PTAs",
                               "Unemployment Difference",
                               "Export (logged), R to D",
                               "Export (logged), D to R",
                               "Other Economic Aid",
                               "Military aid"),
                               #"Democracy * Naturalization",
                               #"Democracy (squared) Naturalization"),
          omit = c("ccode", "19", "20"),
          add.lines = list(
            c(diagnostics[1], round(c(fit.base$r.squared,
                                      fit.base_aid$r.squared,
                                      fit.pre$r.squared,
                                      fit.post$r.squared,
                                      fit.out$r.squared),
                                      2)),
            c(diagnostics[2], round(c(fit.base$fstatistic[1],
                                      fit.base_aid$fstatistic[1],
                                      fit.pre$fstatistic[1],
                                      fit.post$fstatistic[1],
                                      fit.out$fstatistic[1]),
                                      2)),
            c(diagnostics[3], nrow(mergepdata4) - nrow(as.matrix(fit.base$na.action)),
              nrow(mergepdata4) - nrow(as.matrix(fit.base_aid$na.action)),
              nrow(mergepdata4_pre) - nrow(as.matrix(fit.pre$na.action)),
              nrow(mergepdata4_post) - nrow(as.matrix(fit.post$na.action)),
              nrow(mergepdata4_out) - nrow(as.matrix(fit.out$na.action)))))

```


*********************

diagnostics for linear regression
https://data.library.virginia.edu/diagnostic-plots/
https://www.statmethods.net/stats/rdiagnostics.html


*random note on something else
  - good tutorial on roll_sum
    https://itsalocke.com/blog/understanding-rolling-calculations-in-r/
    
```{r}

options(scipen=999)
vif(base_m)
options(scipen=0)
#naturalization is somewhat okay but variables like population and gdp are very high
# https://statisticalhorizons.com/multicollinearity
  # this says that this is ok as long as these variables are controls, which is true in my case
  # multicollinearity affects only variables that are collinear
  #GVIF^(1/(2*Df)) I think is good when DF > 1.
    # https://stats.stackexchange.com/questions/70679/which-variance-inflation-factor-should-i-be-using-textgvif-or-textgvif

qqPlot(base_m, main="QQ Plot")
# to check that the residuals have a normal distribution
  # a qqplot that is mostly a diagonal line supports this 
  #compare between residuals and theoretical quantiles from a theoretical distribution

influencePlot(base_m, id.method="identify", 
              main="Influence Plot", 
              sub="Circle size is proportial to Cook's Distance" )



```



**********


variables to test
```{r}
log_naturalization_roll_df_lag
naturalization_roll_prop_df_lag

```

***


Dynamic model as a robustness check (Might not include this)
```{r}


base_dynamic_m <- lm(democracy_aid_com_corr ~ 
                       democracy_aid_com_corr_lag +
                       ipdiff_5_lag + 
                      r_conflict_prioryear_lag +
                      r_politychange_lag +
                      r_democracy_lag +
                      r_democracy_sq_lag +
                      r_gdp_growth_lag +
                      d_gdp_growth_lag +
                      r_gdplog_lag +
                      d_gdplog_lag +
                      r_population_lag +
                      d_population_lag +
                      d_women_parliament_lag +
                      comlang_off_lag +
                      colony_lag +
                      distlog_lag +
                      ptacount_lag +
                      unempdiff_lag +
                      log_naturalization_lag +
                      log_rtod_export_lag +
                      log_dtor_export_lag +
                      other_economic_aid_lag +
                      military_aid_corr_lag +
                      ccode1 + 
                      ccode2 + 
                      as.factor(year),
                     data = mergepdata4,
                     na.action = na.omit)

summary(base_dynamic_m)


base_dynamic_m_hc3 <- coeftest(base_dynamic_m, cluster.vcov(base_dynamic_m,
                                            cluster = cbind(mergepdata4$ccode1,
                                                            mergepdata4$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(24:nrow(summary(base_dynamic_m)$coefficients)),]; base_dynamic_m_hc3


# naturalization still holds but need to account for the effect of the lagged dependent variable

```



First Differenced Model
```{r}

#first difference
  # this will essentially be fixed effects for dyads

base_fd_lag_m <- lm(democracy_aid_com_corr_df ~ 
                      democracy_aid_com_corr_df_lag +
                      log_naturalization_df_lag +
                      ipfd_5_lag +
                       r_conflict_prioryear_df_lag +
                      r_politychange_df_lag +
                      r_democracy_df_lag +
                      r_democracy_sq_df_lag +
                      r_gdp_growth_df_lag +
                      d_gdp_growth_df_lag +
                      r_gdplog_df_lag +
                      d_gdplog_df_lag +
                      r_population_df_lag +
                      d_population_df_lag +
                      d_women_parliament_df_lag +
                      ptacount_df_lag +
                      unempdiff_df_lag +
                      log_rtod_export_df_lag +
                      log_dtor_export_df_lag +
                      other_economic_aid_df_lag +
                      military_aid_corr_df_lag,
                     data = mergepdata4,
                     na.action = na.omit)


summary(base_fd_lag_m)


base_fd_lag_m_hc3 <- coeftest(base_fd_lag_m, cluster.vcov(base_fd_lag_m,
                                            cluster = cbind(mergepdata4$ccode1,
                                                            mergepdata4$ccode2), 
                                            leverage = 3, 
                                            df_correction = T)); base_fd_lag_m_hc3


#direction is negative when I used log_naturalization_df_lag
#direction is positive (but not significant) when I use log_naturalization_roll_df_lag 


#therefore, log_naturalization_roll has been more consistent 
  #maybe use this instead


  #However, when I try with log_naturalization_lag instead of df lag (because I think naturalization is already fd) it holds

#singular when military aid is added (corrected- there was an issue with the variable)


#direction is negative when I used log_naturalization_df_lag but not significant 
#direction is positive (but not significant) when I use log_naturalization_roll_df_lag 

#therefore, log_naturalization_roll has been more consistent here too


#should dyad FE still be included here?

```


***

Latex table: Dynamic model
```{r}

fit.dynamic <- summary(base_dynamic_m, diagnostics = T)
fit.fd <- summary(base_fd_lag_m, diagnostics = T)
fit.fd_aid <- summary(base_fd_aidlag_m, diagnostics = T)

stargazer(base_dynamic_m_hc3,
          type = "latex",
          covariate.labels = c("Democracy Aid (lagged)",
                               "Ideal Point Diff. (5 year lag)",
                               "Conflict, Prior Year, R",
                               "Democratization (ord), R",
                               "Democracy (ord), R",
                               "Democracy (squared, ord), R",
                               "GDP growth, R",
                               "GDP growth, D",
                               "GDP (logged), R",
                               "GDP (logged), D",
                               "Population (logged), R",
                               "Population (logged), D",
                               "Women in Parliament (%), D",
                               "Common Official Lang.",
                               "Past Colonial Relationship",
                               "Distance (km, logged)",
                               "# of PTAs",
                               "Unemployment Difference",
                               "Naturalization (logged), R to D",
                               "Export (logged), R to D",
                               "Export (logged), D to R"),
          omit = c("ccode", "19", "20"),
          add.lines = list(
            c(diagnostics[1], round(c(fit.dynamic$r.squared),
                                      2)),
            c(diagnostics[2], round(c(fit.dynamic$fstatistic[1]),
                                      2)),
            c(diagnostics[3], nrow(mergepdata4) - nrow(as.matrix(fit.dynamic$na.action)))))

```



Just the first differenced models
```{r}
stargazer(base_fd_lag_m,
          base_fd_aidlag_m,
          type = "latex",
          covariate.labels = c("Democracy Aid (lagged)",
                               "Ideal Point Diff. (5 year lag)",
                               "Conflict, Prior Year, R",
                               #"Democratization (ord), R",
                               "Democracy (ord), R",
                               "Democracy (squared, ord), R",
                               #"GDP growth, R",
                               #"GDP growth, D",
                               "GDP (logged), R",
                               "GDP (logged), D",
                               "Population (logged), R",
                               "Population (logged), D",
                               "Women in Parliament (%), D",
                               #"Common Official Lang.",
                               #"Past Colonial Relationship",
                               #"Distance (km, logged)",
                               "# of PTAs",
                               "Unemployment Difference",
                               "Naturalization (logged), R to D",
                               "Export (logged), R to D",
                               "Export (logged), D to R",
                               "Other Economic Aid"),
                               #"Military Aid"),
          omit = c("ccode", "19", "20"))


```

****



ECM

```{r}


base_ecm_m <- lm(democracy_aid_com_corr_df ~ 
                   democracy_aid_com_corr_lag +
                   log_naturalization_df +
                   log_naturalization_lag +
                   ipfd +
                   ipdiff_lag +
                   r_conflict_prioryear_df +
                   r_conflict_prioryear_lag +
                   r_politychange_df +
                   r_politychange_lag +
                   r_democracy_df +
                   r_democracy_lag +
                   r_democracy_sq_df +
                   r_democracy_sq_lag +
                   r_gdp_growth_df +
                   r_gdp_growth_lag +
                   d_gdp_growth_df +
                   d_gdp_growth_lag +
                   r_gdplog_df +
                   r_gdplog_lag +
                   d_gdplog_df +
                   d_gdplog_lag +
                   r_population_df +
                   r_population_lag +
                   d_population_df +
                   d_population_lag +
                   d_women_parliament_df +
                   d_women_parliament_lag +
                   ptacount_df+
                   ptacount_lag +
                   unempdiff_df +
                   unempdiff_lag +
                   #log_naturalization_roll_df +
                   #log_naturalization_roll_lag +
                   log_rtod_export_df +
                   log_rtod_export_lag +
                   log_dtor_export_df +
                   log_dtor_export_lag +
                   other_economic_aid_df +
                   other_economic_aid_lag +
                   military_aid_corr_df +
                   military_aid_corr_lag +
                   ccode1 +
                   ccode2 +
                   as.factor(year),
                 data = mergepdata4,
                 na.action = na.exclude)


#summary(base_ecm_m)


base_ecm_m_hc3 <- coeftest(base_ecm_m, cluster.vcov(base_ecm_m,
                                            cluster = cbind(mergepdata4$ccode1,
                                                            mergepdata4$ccode2), 
                                            leverage = 3, 
                                            df_correction = T)) #; base_ecm_m_hc3

#effect (at least the short term effect for now, holds for log naturalization)

```




```{r}

#get predicted values
ecm_pred_m1 <- as.data.frame(cbind(names(predict(base_ecm_m)),
                          predict(base_ecm_m),
                          1995:2016)) %>%
  rename(dyad_index = V1,
         pred_ecm_m1 = V2,
         year = V3) %>%
  mutate(dyad_index = gsub("\\..*", "", dyad_index)) %>%
  separate(dyad_index, c("ccode1", "ccode2"), "-") %>%
  mutate(pred_ecm_m1 = as.numeric(as.character(pred_ecm_m1)),
         year = as.numeric(as.character(year)))
  
mergepdata4_ecm <- left_join(mergepdata4,
                             ecm_pred_m1,
                             by = c("ccode1",
                                    "ccode2",
                                    "year")) %>%
  pdata.frame(.)



#Bewley transformation

base_ecm_bewley_m <- lm(democracy_aid_com_corr ~ 
                   pred_ecm_m1 +
                   log_naturalization_df +
                   log_naturalization_lag +
                   ipfd +
                   ipdiff_lag +
                   r_conflict_prioryear_df +
                   r_conflict_prioryear_lag +
                   r_politychange_df +
                   r_politychange_lag +
                   r_democracy_df +
                   r_democracy_lag +
                   r_democracy_sq_df +
                   r_democracy_sq_lag +
                   r_gdp_growth_df +
                   r_gdp_growth_lag +
                   d_gdp_growth_df +
                   d_gdp_growth_lag +
                   r_gdplog_df +
                   r_gdplog_lag +
                   d_gdplog_df +
                   d_gdplog_lag +
                   r_population_df +
                   r_population_lag +
                   d_population_df +
                   d_population_lag +
                   d_women_parliament_df +
                   d_women_parliament_lag +
                   ptacount_df+
                   ptacount_lag +
                   unempdiff_df +
                   unempdiff_lag +
                   #log_naturalization_roll_df +
                   #log_naturalization_roll_lag +
                   log_rtod_export_df +
                   log_rtod_export_lag +
                   log_dtor_export_df +
                   log_dtor_export_lag +
                   other_economic_aid_df +
                   other_economic_aid_lag +
                   military_aid_corr_df +
                   military_aid_corr_lag +
                   ccode1 +
                   ccode2 +
                   as.factor(year),
                     data = mergepdata4_ecm,
                     na.action = na.exclude)


#summary(base_ecm_m)


base_ecm_bewley_m_hc3 <- coeftest(base_ecm_bewley_m, cluster.vcov(base_ecm_bewley_m,
                                            cluster = cbind(mergepdata4_ecm$ccode1,
                                                            mergepdata4_ecm$ccode2), 
                                            leverage = 3, 
                                            df_correction = T)); base_ecm_bewley_m_hc3


```

***

Latex for ECM

```{r}

fit.ecm <- summary(base_ecm_m, diagnostics = T)
fit.ecm_bewley <- summary(base_ecm_bewley_m, diagnostics = T)



stargazer(base_ecm_m_hc3,
          base_ecm_bewley_m_hc3,
          omit = c("ccode", "19", "20", "pred_ecm_m1", "gdp_growth"),
          covariate.labels = c("Lag of Democracy aid (logged, per 1000)",
                               "fd Naturalization (logged), R to D",
                               "Lagged Naturalization (logged), R to D",
                               "fd Ideal Point Diff.",
                               "Lagged Ideal Point Diff.",
                               "fd Conflict, R",
                               "Lagged Conflict, R",
                               "fd Democratization (ord), R",
                               "Lagged Democratization (ord), R",
                               "fd Democracy (ord), R",
                               "Lagged Democracy (ord), R",
                               "fd Democracy (squared, ord), R",
                               "Lagged Democracy (squared, ord), R",
                               #"fd GDP growth, R",
                               #"GDP growth, R",
                               #"fd GDP growth, D",
                               #"GDP growth, D",
                               "fd GDP (logged), R",
                               "Lagged GDP (logged), R",
                               "fd GDP (logged), D",
                               "Lagged GDP (logged), D",
                               "fd Population (logged), R",
                               "Lagged Population (logged), R",
                               "fd Population (logged), D",
                               "Lagged Population (logged), D",
                               "fd Women in Parliament (%), D",
                               "Lagged Women in Parliament (%), D",
                               "fd # of PTAs",
                               "Lagged # of PTAs",
                               "fd Unemployment Difference",
                               "Lagged Unemployment Difference",
                               "fd Export (logged), R to D",
                               "Lagged Export (logged), R to D",
                               "fd Export (logged), D to R",
                               "Lagged Export (logged), D to R",
                               "fd Other Economic Aid (logged)",
                               "Lagged Other Economic Aid",
                               "fd US Military Aid (logged)",
                               "Lagged US Military Aid (logged)"),
          add.lines = list(
            c(diagnostics[1], round(c(fit.ecm$r.squared,
                                      fit.ecm_bewley$r.squared),
                                      2)),
            c(diagnostics[2], round(c(fit.ecm$fstatistic[1],
                                      fit.ecm_bewley$fstatistic[1]),
                                      2)),
            c(diagnostics[3], nrow(mergepdata4) - nrow(as.matrix(fit.ecm$na.action)),
              nrow(mergepdata4) - nrow(as.matrix(fit.ecm_bewley$na.action)))),
          type = "latex")



```





***

Dynamic simulations of the ECM model

```{r}

#low naturalization 
scenario1 <- mergepdata4 %>%
  mutate_at(vars(democracy_aid_com_corr_lag,
                 ipfd_5,
                 ipdiff_5_lag,
                 r_conflict_prioryear_df, 
                 r_conflict_prioryear_lag, 
                 r_politychange_df,
                 r_politychange_lag,
                 r_democracy_df,
                 r_democracy_lag,
                 r_democracy_sq_df,
                 r_democracy_sq_lag, 
                 r_gdp_growth_df, 
                 r_gdp_growth_lag, 
                 d_gdp_growth_df, 
                 d_gdp_growth_lag, 
                 r_gdplog_df, 
                 r_gdplog_lag, 
                 d_gdplog_df,
                 d_gdplog_lag,
                 r_population_df,
                 r_population_lag,
                 d_population_df,
                 d_population_lag,
                 d_women_parliament_df,
                 d_women_parliament_lag,
                 ptacount_df,
                 ptacount_lag,
                 unempdiff_df,
                 unempdiff_lag,
                 log_rtod_export_df,
                 log_rtod_export_lag,
                 log_dtor_export_df,
                 log_dtor_export_lag,
                 other_economic_aid_df,
                 other_economic_aid_lag,
                 military_aid_corr_df,
                 military_aid_corr_lag), 
            mean, na.rm = T) %>%
  mutate(log_naturalization_df = 0,
         log_naturalization_lag = quantile(log_naturalization_lag, 0.05, na.rm = T)) %>%
  select(log_naturalization_df,
         log_naturalization_lag,
         democracy_aid_com_corr_lag,
         ipfd_5,
                 ipdiff_5_lag,
                 r_conflict_prioryear_df, 
                 r_conflict_prioryear_lag, 
                 r_politychange_df,
                 r_politychange_lag,
                 r_democracy_df,
                 r_democracy_lag,
                 r_democracy_sq_df,
                 r_democracy_sq_lag, 
                 r_gdp_growth_df, 
                 r_gdp_growth_lag, 
                 d_gdp_growth_df, 
                 d_gdp_growth_lag, 
                 r_gdplog_df, 
                 r_gdplog_lag, 
                 d_gdplog_df,
                 d_gdplog_lag,
                 r_population_df,
                 r_population_lag,
                 d_population_df,
                 d_population_lag,
                 d_women_parliament_df,
                 d_women_parliament_lag,
                 ptacount_df,
                 ptacount_lag,
                 unempdiff_df,
                 unempdiff_lag,
                 log_naturalization_df,
                 log_naturalization_lag,
                 log_rtod_export_df,
                 log_rtod_export_lag,
                 log_dtor_export_df,
                 log_dtor_export_lag,
                 other_economic_aid_df,
                 other_economic_aid_lag,
                 military_aid_corr_df,
                 military_aid_corr_lag) %>%
  distinct()
  
  


#high naturalization 
scenario2 <- mergepdata4 %>%
  mutate_at(vars(democracy_aid_com_corr_lag,
                 ipfd_5,
                 ipdiff_5_lag,
                 r_conflict_prioryear_df, 
                 r_conflict_prioryear_lag, 
                 r_politychange_df,
                 r_politychange_lag,
                 r_democracy_df,
                 r_democracy_lag,
                 r_democracy_sq_df,
                 r_democracy_sq_lag, 
                 r_gdp_growth_df, 
                 r_gdp_growth_lag, 
                 d_gdp_growth_df, 
                 d_gdp_growth_lag, 
                 r_gdplog_df, 
                 r_gdplog_lag, 
                 d_gdplog_df,
                 d_gdplog_lag,
                 r_population_df,
                 r_population_lag,
                 d_population_df,
                 d_population_lag,
                 d_women_parliament_df,
                 d_women_parliament_lag,
                 ptacount_df,
                 ptacount_lag,
                 unempdiff_df,
                 unempdiff_lag,
                 log_rtod_export_df,
                 log_rtod_export_lag,
                 log_dtor_export_df,
                 log_dtor_export_lag,
                 other_economic_aid_df,
                 other_economic_aid_lag,
                 military_aid_corr_df,
                 military_aid_corr_lag), 
            mean, na.rm = T) %>%
  mutate(log_naturalization_df = 0,
         log_naturalization_lag = quantile(log_naturalization_lag, 0.95, na.rm = T)) %>%
  select(log_naturalization_df,
         log_naturalization_lag,
         democracy_aid_com_corr_lag,
         ipfd_5,
                 ipdiff_5_lag,
                 r_conflict_prioryear_df, 
                 r_conflict_prioryear_lag, 
                 r_politychange_df,
                 r_politychange_lag,
                 r_democracy_df,
                 r_democracy_lag,
                 r_democracy_sq_df,
                 r_democracy_sq_lag, 
                 r_gdp_growth_df, 
                 r_gdp_growth_lag, 
                 d_gdp_growth_df, 
                 d_gdp_growth_lag, 
                 r_gdplog_df, 
                 r_gdplog_lag, 
                 d_gdplog_df,
                 d_gdplog_lag,
                 r_population_df,
                 r_population_lag,
                 d_population_df,
                 d_population_lag,
                 d_women_parliament_df,
                 d_women_parliament_lag,
                 ptacount_df,
                 ptacount_lag,
                 unempdiff_df,
                 unempdiff_lag,
                 log_naturalization_df,
                 log_naturalization_lag,
                 log_rtod_export_df,
                 log_rtod_export_lag,
                 log_dtor_export_df,
                 log_dtor_export_lag,
                 other_economic_aid_df,
                 other_economic_aid_lag,
                 military_aid_corr_df,
                 military_aid_corr_lag) %>%
  distinct()

```

Attempt to simulate Dynamic model
```{r}

#dynamic model



#low naturalization 
dynscenario1 <- mergepdata4 %>%
  mutate_at(vars(democracy_aid_com_corr_df,
                 democracy_aid_com_corr_df_lag,
                      ipfd_5_lag,
                       r_conflict_prioryear_df_lag,
                      r_politychange_df_lag,
                      r_democracy_df_lag,
                      r_democracy_sq_df_lag,
                      r_gdp_growth_df_lag,
                      d_gdp_growth_df_lag,
                      r_gdplog_df_lag,
                      d_gdplog_df_lag,
                      r_population_df_lag,
                      d_population_df_lag,
                      d_women_parliament_df_lag,
                      ptacount_df_lag,
                      unempdiff_df_lag,
                      log_rtod_export_df_lag,
                      log_dtor_export_df_lag,
                      other_economic_aid_df_lag,
                      military_aid_corr_df_lag), 
            mean, na.rm = T) %>%
  mutate(log_naturalization_df_lag = quantile(log_naturalization_df_lag, 0.05, na.rm = T)) %>%
  select(log_naturalization_df_lag,
         democracy_aid_com_corr_df,
                 democracy_aid_com_corr_df_lag,
                      log_naturalization_df_lag,
                      ipfd_5_lag,
                       r_conflict_prioryear_df_lag,
                      r_politychange_df_lag,
                      r_democracy_df_lag,
                      r_democracy_sq_df_lag,
                      r_gdp_growth_df_lag,
                      d_gdp_growth_df_lag,
                      r_gdplog_df_lag,
                      d_gdplog_df_lag,
                      r_population_df_lag,
                      d_population_df_lag,
                      d_women_parliament_df_lag,
                      ptacount_df_lag,
                      unempdiff_df_lag,
                      log_rtod_export_df_lag,
                      log_dtor_export_df_lag,
                      other_economic_aid_df_lag,
                      military_aid_corr_df_lag) %>%
  distinct()
  
  

#high naturalization 

dynscenario2 <- mergepdata4 %>%
  mutate_at(vars(democracy_aid_com_corr_df,
                 democracy_aid_com_corr_df_lag,
                      ipfd_5_lag,
                       r_conflict_prioryear_df_lag,
                      r_politychange_df_lag,
                      r_democracy_df_lag,
                      r_democracy_sq_df_lag,
                      r_gdp_growth_df_lag,
                      d_gdp_growth_df_lag,
                      r_gdplog_df_lag,
                      d_gdplog_df_lag,
                      r_population_df_lag,
                      d_population_df_lag,
                      d_women_parliament_df_lag,
                      ptacount_df_lag,
                      unempdiff_df_lag,
                      log_rtod_export_df_lag,
                      log_dtor_export_df_lag,
                      other_economic_aid_df_lag,
                      military_aid_corr_df_lag), 
            mean, na.rm = T) %>%
  mutate(log_naturalization_df_lag = quantile(log_naturalization_df_lag, 0.95, na.rm = T)) %>%
  select(log_naturalization_df_lag,
         democracy_aid_com_corr_df,
                 democracy_aid_com_corr_df_lag,
                      log_naturalization_df_lag,
                      ipfd_5_lag,
                       r_conflict_prioryear_df_lag,
                      r_politychange_df_lag,
                      r_democracy_df_lag,
                      r_democracy_sq_df_lag,
                      r_gdp_growth_df_lag,
                      d_gdp_growth_df_lag,
                      r_gdplog_df_lag,
                      d_gdplog_df_lag,
                      r_population_df_lag,
                      d_population_df_lag,
                      d_women_parliament_df_lag,
                      ptacount_df_lag,
                      unempdiff_df_lag,
                      log_rtod_export_df_lag,
                      log_dtor_export_df_lag,
                      other_economic_aid_df_lag,
                      military_aid_corr_df_lag) %>%
  distinct()
  


scenes <- list(scenario1, scenario2)
dynscenes <- list(dynscenario1, dynscenario2)


```


```{r}


#does not run for fd model
summary(base_fd_lag_m)


aa <- dynsim(obj = base_ecm_m, ldv = "democracy_aid_com_corr_lag", scen = scenes)

dynsimGG(dynamic_sim, leg.name = "Scenarios", leg.labels = c("5th Percentile", "95th Percentile"), 
          #color = "OrRd",
          ylab = "Predicted Democracy Aid Allocation")

#there is a diff, but it overlaps and is consistent over time
#maybe do this with a lagged model?
  #looking at greene's do file and best's explanation I think the method (doing it on ECM) correct
    #however, I think they make the fd of the independent variable as 0 - makes sense, since t-1 is constant
    #made df 0, but it is still the same
#maybe this is better for interaction?
  #I don't think it should make a dif


```




***
case-study selection

```{r}

#lieberman 2005 says that if stats are robust, case selection should be made on the line
  #low residual cases

#let's just squish everything to means
class(mergepdata4_ecm$pred_ecm_m1)

case_sel <- mergepdata4 %>%
  group_by(ccode1, ccode2) %>%
   select(log_naturalization_df,
         log_naturalization_lag,
         democracy_aid_com_corr_lag,
         democracy_aid_com_corr_df,
         ipfd_5,
                 ipdiff_5_lag,
                 r_conflict_prioryear_df, 
                 r_conflict_prioryear_lag, 
                 r_politychange_df,
                 r_politychange_lag,
                 r_democracy_df,
                 r_democracy_lag,
                 r_democracy_sq_df,
                 r_democracy_sq_lag, 
                 r_gdp_growth_df, 
                 r_gdp_growth_lag, 
                 d_gdp_growth_df, 
                 d_gdp_growth_lag, 
                 r_gdplog_df, 
                 r_gdplog_lag, 
                 d_gdplog_df,
                 d_gdplog_lag,
                 r_population_df,
                 r_population_lag,
                 d_population_df,
                 d_population_lag,
                 d_women_parliament_df,
                 d_women_parliament_lag,
                 ptacount_df,
                 ptacount_lag,
                 unempdiff_df,
                 unempdiff_lag,
         other_economic_aid_df,
         other_economic_aid_lag,
                 log_naturalization_df,
                 log_naturalization_lag,
                 log_rtod_export_df,
                 log_rtod_export_lag,
                 log_dtor_export_df,
                 log_dtor_export_lag) %>%
  summarise_all(mean, na.rm = T)


base_ecm_case_m <- lm(democracy_aid_com_corr_df ~ 
                   democracy_aid_com_corr_lag +
                   ipfd_5 +
                   ipdiff_5_lag +
                   r_conflict_prioryear_df +
                   r_conflict_prioryear_lag +
                   r_politychange_df +
                   r_politychange_lag +
                   r_democracy_df +
                   r_democracy_lag +
                   r_democracy_sq_df +
                   r_democracy_sq_lag +
                   r_gdp_growth_df +
                   r_gdp_growth_lag +
                   d_gdp_growth_df +
                   d_gdp_growth_lag +
                   r_gdplog_df +
                   r_gdplog_lag +
                   d_gdplog_df +
                   d_gdplog_lag +
                   r_population_df +
                   r_population_lag +
                   d_population_df +
                   d_population_lag +
                   d_women_parliament_df +
                   d_women_parliament_lag +
                   ptacount_df+
                   ptacount_lag +
                   unempdiff_df +
                   unempdiff_lag +
                   #log_naturalization_roll_df +
                   #log_naturalization_roll_lag +
                   log_naturalization_df +
                   log_naturalization_lag +
                   log_rtod_export_df +
                   log_rtod_export_lag +
                   log_dtor_export_df +
                   log_dtor_export_lag +
                   other_economic_aid_df +
                   other_economic_aid_lag +
                   ccode1 +
                   ccode2,
                   #as.factor(year),
                   #military_aid_df +
                   #military_aid_lag,
                 data = case_sel,
                 na.action = na.exclude)


case_sel$pred <- predict(base_ecm_case_m)
case_sel$res <- resid(base_ecm_case_m)

case_sel_simple <- case_sel %>%
  select(ccode1,
         ccode2,
         democracy_aid_com_corr_df,
         log_naturalization_lag,
         pred,
         res) %>%
  mutate(res_abs = abs(res),
         donor = countrycode(ccode1, "cown", "country.name"),
         recipient = countrycode(ccode2, "cown", "country.name"))


#get predicted values

ggplot(case_sel, aes(x=pred, y=democracy_aid_com_corr_df)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color="red")


 
```
 
```{r}
 ###
 
 
 #netherlands (210) and indonesia (850)
  #idn was a colony of netherlands
  #during national revolution (1945-1949) hundreds of thousands indonesians moved to netherlands
  #wiki says that there are 700 000 1st and 2nd gen indonesians in netherlands
  #netherland and indonesia are big donors and recipients
  #in 1999, idn had the first election after a regime change
  #1999 was the point that recorded the most indonesians naturalizing in netherlands (500)
    #and 2000 was the point that received the most democracy aid from netherlands
  #2002 amendments to constitution were seen as important steps to full democratization
    #therefore it might be that the 1999 elections signalled dutch donors that democratization was happending and it just happened to coincide with naturalization, or a combination of both
    #there was also another peak in 2002-2003 that was separate 
  #also, remember that most immigrants were actually already dutch citizen
 
 #however, overal migration from indonesia had really diminished after 1940s
 
 #however, bosma (2012) says that indonesian diaspora were motivated by transnationalism and identity
 
#maybe it is not the naturalized citizen numbers that have an immediate impact? maybe they revive national movements amongst 2nd gen and older diaspora? - since max idn naturalization was only 500

 
 nld_idn <- mergedata4 %>%
   filter(dyad_id == "210_850")
 

ggplot(nld_idn, aes(x = year,
                    y = naturalization)) +
  geom_line()

ggplot(nld_idn, aes(x = year,
                    y = democracy_aid_com)) +
  geom_line()
 

ggplot(mergedata4, aes(log_naturalization)) + 
  geom_histogram()


#lets also look at surinamese (115) in nld
#bosma 2012 shows that surinamese organizations opened up the most in nld
nld_sur <- mergedata4 %>%
   filter(dyad_id == "210_115")

ggplot(nld_sur, aes(x = year,
                    y = naturalization)) +
  geom_line()
ggplot(nld_sur, aes(x = year,
                    y = democracy_aid_com)) +
  geom_line()
 


#UK(200) and india (750)
uk_ind <- mergedata4 %>%
   filter(dyad_id == "200_750")

ggplot(uk_ind, aes(x = year,
                    y = naturalization)) +
  geom_line()
ggplot(uk_ind, aes(x = year,
                    y = democracy_aid_com)) +
  geom_line()

#a lot of democracy aid - probably due to population scales
#although it somewhat seems like naturalization is following aid

#US(2) and burma (775)
us_bur <- mergedata4 %>%
   filter(dyad_id == "2_775")

ggplot(us_bur, aes(x = year,
                    y = naturalization)) +
  geom_line()
ggplot(us_bur, aes(x = year,
                    y = democracy_aid_com)) +
  geom_line()


#US(2) and cuba (40)
us_cub <- mergedata4 %>%
   filter(dyad_id == "2_40")

ggplot(us_cub, aes(x = year,
                   y = naturalization)) +
  geom_line()
ggplot(us_cub, aes(x = year,
                   y = democracy_aid_com)) +
  geom_line()


#korea(732) and thailand(800)
kor_tha <- mergedata4 %>%
   filter(dyad_id == "732_800")

ggplot(kor_tha, aes(x = year,
                   y = naturalization)) +
  geom_line()

ggplot(kor_tha, aes(x = year,
                   y = democracy_aid_com)) +
  geom_line()



#korea(732) and nepal(790)
kor_nep <- mergedata4 %>%
   filter(dyad_id == "732_790")

ggplot(kor_nep, aes(x = year,
                   y = naturalization)) +
  geom_line()

ggplot(kor_nep, aes(x = year,
                   y = democracy_aid_com)) +
  geom_line()



#korea(732) and vietnam(817)
kor_vie <- mergedata4 %>%
   filter(dyad_id == "732_817")

ggplot(kor_vie, aes(x = year,
                   y = naturalization)) +
  geom_line()

ggplot(kor_vie, aes(x = year,
                   y = democracy_aid_com)) +
  geom_line()

#spike at 2011, which was when the national assembly elections were held in vietnam


#korea(732) and vietnam(817)
kor_vie <- mergedata4 %>%
   filter(dyad_id == "732_817")

ggplot(kor_vie, aes(x = year,
                   y = naturalization)) +
  geom_line()

ggplot(kor_vie, aes(x = year,
                   y = democracy_aid_com)) +
  geom_line()

#spike at 2011, which was when the national assembly elections were held in vietnam


```




*****
*****

Appendix

***
Base Model - with disbursement
```{r}
#all covariates lagged 1

base_disb_m <- lm(democracy_aid_disb_corr ~ 
                ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag +
                r_democracy_sq_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                log_naturalization_lag +
                #log_immigration_lag +
                #naturalization_prop_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4,
              na.action = na.omit)

#summary(base_m)


base_disb_m_hc3 <- coeftest(base_disb_m, cluster.vcov(base_disb_m,
                                            cluster = cbind(mergepdata4$ccode1, mergepdata4$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(22:nrow(summary(base_disb_m)$coefficients)),]; base_disb_m_hc3


#with disbursement, naturalization still holds




```

Disbursement with other aid
```{r}
#base model with other aid types - disbursement

base_disb_m_aid <- lm(democracy_aid_disb_corr ~ 
                   ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag +
                r_democracy_sq_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                log_naturalization_lag +
                #log_immigration_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                other_economic_aid_lag +
                military_aid_corr_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4,
              na.action = na.omit)

summary(base_disb_m_aid)

#all aid except for military aid affects democracy aid (positively)

base_disb_m_aid_hc3 <- coeftest(base_disb_m_aid, cluster.vcov(base_disb_m_aid,
                                            cluster = cbind(mergepdata4$ccode1, mergepdata4$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(24:nrow(summary(base_disb_m_aid)$coefficients)),]; base_disb_m_aid_hc3

#with disbursement, naturalization still holds

```


***
Base model with aid using prop of naturalization

```{r}


base_prop_m_aid <- lm(democracy_aid_disb_corr ~ 
                   ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag +
                r_democracy_sq_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                #log_naturalization_lag +
                #log_immigration_lag +
                  naturalization_prop_log_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                other_economic_aid_lag +
                military_aid_corr_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4,
              na.action = na.omit)

#summary(base_prop_m_aid)

#all aid except for military aid affects democracy aid (positively)

base_prop_m_aid_hc3 <- coeftest(base_prop_m_aid, cluster.vcov(base_prop_m_aid,
                                            cluster = cbind(mergepdata4$ccode1, mergepdata4$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(24:nrow(summary(base_prop_m_aid)$coefficients)),]; base_prop_m_aid_hc3

#effect of naturalization holds when it is scaled to donor country pop


```


Base model using roll sum

```{r}


base_roll_m_aid <- lm(democracy_aid_disb_corr ~ 
                   ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag +
                r_democracy_sq_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                #log_naturalization_lag +
                #log_immigration_lag +
                #naturalization_prop_lag +
                  log_naturalization_roll_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                other_economic_aid_lag +
                military_aid_corr_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4,
              na.action = na.omit)

#summary(base_prop_m_aid)

#all aid except for military aid affects democracy aid (positively)

base_roll_m_aid_hc3 <- coeftest(base_roll_m_aid, cluster.vcov(base_roll_m_aid,
                                            cluster = cbind(mergepdata4$ccode1, mergepdata4$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(24:nrow(summary(base_roll_m_aid)$coefficients)),]; base_roll_m_aid_hc3

#when proportional to population using roll sum, the effect holds too


```

***

Base model using roll sum prop

```{r}


base_roll_prop_m_aid <- lm(democracy_aid_disb_corr ~ 
                   ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag +
                r_democracy_sq_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                #log_naturalization_lag +
                #log_immigration_lag +
                #naturalization_prop_lag +
                naturalization_roll_prop_log_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                other_economic_aid_lag +
                military_aid_corr_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4,
              na.action = na.omit)

#summary(base_roll_prop_m_aid)

#all aid except for military aid affects democracy aid (positively)

base_roll_prop_m_aid_hc3 <- coeftest(base_roll_prop_m_aid, cluster.vcov(base_roll_prop_m_aid,
                                            cluster = cbind(mergepdata4$ccode1, mergepdata4$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(24:nrow(summary(base_roll_prop_m_aid)$coefficients)),]; base_roll_prop_m_aid_hc3

#when proportional to population using roll sum, the effect holds too


```

***
Latex
```{r}

fit.disb <- summary(base_disb_m_aid, diagnostics = T)
fit.prop <- summary(base_prop_m_aid, diagnostics = T)
fit.roll <- summary(base_roll_m_aid, diagnostics = T)
fit.rollprop <- summary(base_roll_prop_m_aid, diagnostics = T)





stargazer(base_m_aid_hc3,
          base_disb_m_aid_hc3,
          base_prop_m_aid_hc3,
          base_roll_m_aid_hc3,
          base_roll_prop_m_aid_hc3,
          omit = c("ccode", "19", "20"),
          covariate.labels = c("Ideal Point Diff. (5 year lag)",
                               "Conflict, Prior Year, R",
                               "Democratization (ord), R",
                               "Democracy (ord), R",
                               "Democracy (squared, ord), R",
                               "GDP growth, R",
                               "GDP growth, D",
                               "GDP (logged), R",
                               "GDP (logged), D",
                               "Population (logged), R",
                               "Population (logged), D",
                               "Women in Parliament (%), D",
                               "Common Official Lang.",
                               "Past Colonial Relationship",
                               "Distance (km, logged)",
                               "# of PTAs",
                               "Unemployment Difference",
                               "Naturalization (logged), R to D",
                               "Naturalization (scaled, logged), R to D",
                               "Sum of Naturalization (logged), R to D",
                               "Sum of Naturalization (scaled, logged), R to D",
                               "Export (logged), R to D",
                               "Export (logged), D to R",
                               "Other Economic Aid",
                               "Military aid"),
          add.lines = list(
            c(diagnostics[1], round(c(fit.base$r.squared,
                                      fit.disb$r.squared,
                                      fit.prop$r.squared,
                                      fit.roll$r.squared,
                                      fit.rollprop$r.squared),
                                      2)),
            c(diagnostics[2], round(c(fit.base$fstatistic[1],
                                      fit.disb$fstatistic[1],
                                      fit.prop$fstatistic[1],
                                      fit.roll$fstatistic[1],
                                      fit.rollprop$fstatistic[1]),
                                      2)),
            c(diagnostics[3], nrow(mergepdata4) - nrow(as.matrix(fit.base$na.action)),
              nrow(mergepdata4) - nrow(as.matrix(fit.disb$na.action)),
              nrow(mergepdata4) - nrow(as.matrix(fit.prop$na.action)),
              nrow(mergepdata4) - nrow(as.matrix(fit.roll$na.action)),
              nrow(mergepdata4) - nrow(as.matrix(fit.rollprop$na.action)))),
          type = "latex")



```


***

Visually showing that there aren't a lot of time trend
```{r}
set.seed(1)
sample_dyads <- mergedata4 %>% 
  filter(year > 1994) %>%
  ungroup() %>%
  select(democracy_aid_com_corr, 
         year, 
         dyad_id,
         ccode1,
         ccode2) %>%
  group_by(dyad_id) %>%
  filter(any(democracy_aid_com_corr != 0)) %>%
  ungroup() %>%
  filter(dyad_id %in% sample(dyad_id, 20))%>%
  mutate(dyad_name = paste0(countrycode(ccode1, 
                                        "cown",
                                        "country.name"),
                            " - ",
                            countrycode(ccode2,
                                        "cown",
                                        "country.name")))
  

ggplot(sample_dyads,
       aes(x=year, y=democracy_aid_com_corr)) +
  geom_point() +
  geom_line() +
  facet_wrap(~dyad_name) +
  labs(title = "Democracy aid allocation 1995-2014: Sample of dyads",
       y = "Democracy aid commitment (logged, per 1000 people)",
       x = "Year") +
  scale_x_continuous(breaks = c(2000, 2010))

ggsave("sample_dyad_trends.png", width = 8, height = 7)


#plot makes it seem like there is no stationarity

```


***

Maybe - table with outliers excluded? (Menard 2018 did this as robustness check)
  This is an actual concern because we do have outliers





***
ECM with fungible aid as dependent variable
- program aid


ECM

```{r}


base_ecm_m_program <- lm(program_aid_com_corr_df ~ 
                   program_aid_com_corr_lag +
                   log_naturalization_df +
                   log_naturalization_lag +
                   ipfd +
                   ipdiff_lag +
                   r_conflict_prioryear_df +
                   r_conflict_prioryear_lag +
                   r_politychange_df +
                   r_politychange_lag +
                   r_democracy_df +
                   r_democracy_lag +
                   r_democracy_sq_df +
                   r_democracy_sq_lag +
                   r_gdp_growth_df +
                   r_gdp_growth_lag +
                   d_gdp_growth_df +
                   d_gdp_growth_lag +
                   r_gdplog_df +
                   r_gdplog_lag +
                   d_gdplog_df +
                   d_gdplog_lag +
                   r_population_df +
                   r_population_lag +
                   d_population_df +
                   d_population_lag +
                   d_women_parliament_df +
                   d_women_parliament_lag +
                   ptacount_df+
                   ptacount_lag +
                   unempdiff_df +
                   unempdiff_lag +
                   #log_naturalization_roll_df +
                   #log_naturalization_roll_lag +
                   log_rtod_export_df +
                   log_rtod_export_lag +
                   log_dtor_export_df +
                   log_dtor_export_lag +
                   #other_economic_aid_df +
                   #other_economic_aid_lag +
                   military_aid_corr_df +
                   military_aid_corr_lag +
                   ccode1 +
                   ccode2 +
                   as.factor(year),
                 data = mergepdata4,
                 na.action = na.exclude)


#summary(base_ecm_m)


base_ecm_m_program_hc3 <- coeftest(base_ecm_m_program, cluster.vcov(base_ecm_m_program,
                                            cluster = cbind(mergepdata4$ccode1,
                                                            mergepdata4$ccode2), 
                                            leverage = 3, 
                                            df_correction = T)); base_ecm_m_program_hc3

#effect (at least the short term effect for now, holds for log naturalization)

```




```{r}

#get predicted values
ecm_pred_m1 <- as.data.frame(cbind(names(predict(base_ecm_m)),
                          predict(base_ecm_m),
                          1995:2016)) %>%
  rename(dyad_index = V1,
         pred_ecm_m1 = V2,
         year = V3) %>%
  mutate(dyad_index = gsub("\\..*", "", dyad_index)) %>%
  separate(dyad_index, c("ccode1", "ccode2"), "-") %>%
  mutate(pred_ecm_m1 = as.numeric(as.character(pred_ecm_m1)),
         year = as.numeric(as.character(year)))
  
mergepdata4_ecm <- left_join(mergepdata4,
                             ecm_pred_m1,
                             by = c("ccode1",
                                    "ccode2",
                                    "year")) %>%
  pdata.frame(.)



#Bewley transformation

base_ecm_bewley_m <- lm(democracy_aid_com_corr ~ 
                   pred_ecm_m1 +
                   log_naturalization_df +
                   log_naturalization_lag +
                   ipfd +
                   ipdiff_lag +
                   r_conflict_prioryear_df +
                   r_conflict_prioryear_lag +
                   r_politychange_df +
                   r_politychange_lag +
                   r_democracy_df +
                   r_democracy_lag +
                   r_democracy_sq_df +
                   r_democracy_sq_lag +
                   r_gdp_growth_df +
                   r_gdp_growth_lag +
                   d_gdp_growth_df +
                   d_gdp_growth_lag +
                   r_gdplog_df +
                   r_gdplog_lag +
                   d_gdplog_df +
                   d_gdplog_lag +
                   r_population_df +
                   r_population_lag +
                   d_population_df +
                   d_population_lag +
                   d_women_parliament_df +
                   d_women_parliament_lag +
                   ptacount_df+
                   ptacount_lag +
                   unempdiff_df +
                   unempdiff_lag +
                   #log_naturalization_roll_df +
                   #log_naturalization_roll_lag +
                   log_rtod_export_df +
                   log_rtod_export_lag +
                   log_dtor_export_df +
                   log_dtor_export_lag +
                   other_economic_aid_df +
                   other_economic_aid_lag +
                   military_aid_corr_df +
                   military_aid_corr_lag +
                   ccode1 +
                   ccode2 +
                   as.factor(year),
                     data = mergepdata4_ecm,
                     na.action = na.exclude)


#summary(base_ecm_m)


base_ecm_bewley_m_hc3 <- coeftest(base_ecm_bewley_m, cluster.vcov(base_ecm_bewley_m,
                                            cluster = cbind(mergepdata4_ecm$ccode1,
                                                            mergepdata4_ecm$ccode2), 
                                            leverage = 3, 
                                            df_correction = T)); base_ecm_bewley_m_hc3


```

***

Latex for ECM

```{r}

fit.ecm <- summary(base_ecm_m, diagnostics = T)
fit.ecm_bewley <- summary(base_ecm_bewley_m, diagnostics = T)



stargazer(base_ecm_m_hc3,
          base_ecm_bewley_m_hc3,
          omit = c("ccode", "19", "20", "pred_ecm_m1", "gdp_growth"),
          covariate.labels = c("Lag of Democracy aid (logged, per 1000)",
                               "fd Naturalization (logged), R to D",
                               "Lagged Naturalization (logged), R to D",
                               "fd Ideal Point Diff.",
                               "Lagged Ideal Point Diff.",
                               "fd Conflict, R",
                               "Lagged Conflict, R",
                               "fd Democratization (ord), R",
                               "Lagged Democratization (ord), R",
                               "fd Democracy (ord), R",
                               "Lagged Democracy (ord), R",
                               "fd Democracy (squared, ord), R",
                               "Lagged Democracy (squared, ord), R",
                               #"fd GDP growth, R",
                               #"GDP growth, R",
                               #"fd GDP growth, D",
                               #"GDP growth, D",
                               "fd GDP (logged), R",
                               "Lagged GDP (logged), R",
                               "fd GDP (logged), D",
                               "Lagged GDP (logged), D",
                               "fd Population (logged), R",
                               "Lagged Population (logged), R",
                               "fd Population (logged), D",
                               "Lagged Population (logged), D",
                               "fd Women in Parliament (%), D",
                               "Lagged Women in Parliament (%), D",
                               "fd # of PTAs",
                               "Lagged # of PTAs",
                               "fd Unemployment Difference",
                               "Lagged Unemployment Difference",
                               "fd Export (logged), R to D",
                               "Lagged Export (logged), R to D",
                               "fd Export (logged), D to R",
                               "Lagged Export (logged), D to R",
                               "fd Other Economic Aid (logged)",
                               "Lagged Other Economic Aid",
                               "fd US Military Aid (logged)",
                               "Lagged US Military Aid (logged)"),
          add.lines = list(
            c(diagnostics[1], round(c(fit.ecm$r.squared,
                                      fit.ecm_bewley$r.squared),
                                      2)),
            c(diagnostics[2], round(c(fit.ecm$fstatistic[1],
                                      fit.ecm_bewley$fstatistic[1]),
                                      2)),
            c(diagnostics[3], nrow(mergepdata4) - nrow(as.matrix(fit.ecm$na.action)),
              nrow(mergepdata4) - nrow(as.matrix(fit.ecm_bewley$na.action)))),
          type = "latex")



```









****
Things I'm not sure about including
****

I don't know if I should include this:

I should also test the effect of immigration on naturalization - This is something I can discuss later in the paper as an aside

```{r}

immigration_m <- lm(log_naturalization ~ 
                ipdiff_5 + 
                ipfd_5+
                democracy_aid_disb_corr_lag +
                r_conflict_prioryear +
                r_politychange +
                r_democracy +
                r_gdp_growth +
                d_gdp_growth +
                r_gdplog +
                d_gdplog +
                r_population +
                d_population +
                d_women_parliament +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                r_unemployment_lag +
                d_unemployment_lag +
                unempdiff_1 +
                log_immigration_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata,
              na.action = na.omit)

summary(immigration_m)


immigration_m_hc3 <- coeftest(base_m, cluster.vcov(immigration_m,
                                            cluster = cbind(mergepdata$ccode1, mergepdata$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))[-(24:nrow(summary(immigration_m)$coefficients)),]; immigration_m_hc3


#immigration is a strong predictor of naturalization
  # I think the fact that naturalized citizens can invite others to immigration is a factor
  # if so, it is an omitted variable (nat -> imm -> aid to stop immigration)

#let's recheck immigration and naturalization data to make sure all the lags and stuff are ok

```





****

Maybe I should just look at this using democracy as linear, if I include it
Interaction model
```{r}


int_model <- lm(democracy_aid_com_corr ~ 
                   ipdiff_5 + 
                #ipfd_5+
                r_conflict_prioryear +
                r_politychange_lag +
                r_democracy_lag*log_naturalization_lag +
                r_democracy_sq_lag*log_naturalization_lag +
                r_gdp_growth_lag +
                d_gdp_growth_lag +
                r_gdplog_lag +
                d_gdplog_lag +
                r_population_lag +
                d_population_lag +
                d_women_parliament_lag +
                comlang_off+
                colony +
                distlog+
                ptacount_lag +
                unempdiff_lag +
                #log_immigration_lag +
                log_rtod_export_lag +
                log_dtor_export_lag +
                other_economic_aid_lag +
                military_aid_corr_lag +
                ccode1 + 
                ccode2 + 
                as.factor(year), 
              data = mergepdata4,
              na.action = na.omit)


summary(int_model)
#interesting results
  #conflict and naturalization separately are positive and significant but the interaction is negative and sig

int_model_hc3 <- coeftest(int_model, cluster.vcov(int_model,
                                            cluster = cbind(mergepdata$ccode1, mergepdata$ccode2), 
                                            leverage = 3, 
                                            df_correction = T))#[-(24:nrow(summary(int_model)$coefficients)-2),]; int_model_hc3

#interaction is no longer significant, although it is negative
#interaction with conflict (rare) and democratization (rare) and democracy

### it is significant when we also include democracy ^2

```


```{r}


# interact_plot(int_model, pred = "r_democracy_sq_lag", modx = "log_naturalization_1", data = mergepdata4,
#               linearity.check = TRUE, plot.points = TRUE)


png(filename="interaction.png",
    units="in",
    width=9,
    height=9,
    res=144)

par(mfrow = c(2,2))

interaction_plot_continuous(int_model, 
                            moderator = "log_naturalization_1", 
                            effect = "r_democracy_lag", 
                            "r_democracy_lag:log_naturalization_1",
                            varcov = cluster.vcov(int_model,
                                            cluster = cbind(mergepdata$ccode1, mergepdata$ccode2), 
                                            leverage = 3, 
                                            df_correction = T),
                            xlabel="Naturalization (logged)",
                            ylab = "d(Democracy Aid)/d(Democracy)",
                            title = "")

interaction_plot_continuous(int_model, 
                            effect = "log_naturalization_1", 
                            moderator = "r_democracy_lag", 
                            "r_democracy_lag:log_naturalization_1",
                            varcov = cluster.vcov(int_model,
                                            cluster = cbind(mergepdata$ccode1, mergepdata$ccode2), 
                                            leverage = 3, 
                                            df_correction = T),
                            xlabel="Democracy",
                            ylab = "d(Democracy Aid)/d(Naturalization)",
                            title = "")


interaction_plot_continuous(int_model, 
                            moderator = "log_naturalization_1", 
                            effect = "r_democracy_sq_lag", 
                            "r_democracy_lag:log_naturalization_1",
                            varcov = cluster.vcov(int_model,
                                            cluster = cbind(mergepdata$ccode1, mergepdata$ccode2), 
                                            leverage = 3, 
                                            df_correction = T),
                            xlabel="Naturalization (logged)",
                            ylab = "d(Democracy Aid)/d(Democracy (squared))",
                            title = "")

interaction_plot_continuous(int_model, 
                            effect = "log_naturalization_1", 
                            moderator = "r_democracy_sq_lag", 
                            "r_democracy_lag:log_naturalization_1",
                            varcov = cluster.vcov(int_model,
                                            cluster = cbind(mergepdata$ccode1, mergepdata$ccode2), 
                                            leverage = 3, 
                                            df_correction = T),
                            xlabel="Democracy (squared)",
                            ylab = "d(Democracy Aid)/d(Naturalization)",
                            title = "")

dev.off()

#meplot from https://rpubs.com/milesdwilliams15/381372

# meplot(int_model, 
#        var1 = "log_naturalization_roll_1", 
#        var2 = "r_democracy", 
#        "r_democracy:log_naturalization_roll_1",
#        vcov = cluster.vcov(int_model,
#                              cluster = cbind(mergepdata$ccode1, mergepdata$ccode2), 
#                              leverage = 3, 
#                              df_correction = T))


#like burgess suggests, naturalization*democracy don't really affect one another, but the effect of immigration grows as democracy increases when we have immigration*democracy
#however, then this is not through activism but through direct effect on home country (expat voting)


#** once I include dem^2 it has an effect and is as expected
  #constitutive terms (just democracy related terms) are not significant and only the interactions are


```



****

Dickey-Fuller test

```{r}

# mergepdata_adf <- mergedata3 %>%
#   filter(!is.na(democracy_aid_disb_corr)) %>%
#     pdata.frame(., index = c("dyad_id", "year"))
# 
# mergedata_adf <- mergedata3 %>%
#   filter(!is.na(democracy_aid_disb_corr))

adf_mat <- mergedata3 %>% 
  group_by(dyad_id, year) %>%
  select(democracy_aid_com_corr) %>%
  filter(!is.na(democracy_aid_com_corr)) %>%
  spread(dyad_id, democracy_aid_com_corr) %>%
  select(-no_aid_dyad) %>%
  filter(year > 1995)

missing_dyad <- colnames(adf_mat)[colSums(is.na(adf_mat)) > 0] #none are missing
no_aid_dyad <- colnames(adf_mat)[colSums(adf_mat) < 0.5]
#there is no na, nan, or inf in aa
#also tried removing dyads where no aid was given

purtest(adf_mat[ , sample(2:2160, 10)], 
        pmax = 5, 
        exo = "trend", 
        test = "madwu",
        lags = "AIC") 


#tests: levinlin, ips, madwu, Pm, invnormal, logit, hadri

#levinlin gives an error,
#madwu and others runs buts gives no results
#however, when I limit the number of columns all tests work
  #i think its because some pairs (columns) are problematics and smaller n reduces the likelihood of those columns being picked
  # it worked once with madwu for 100 columns and there was no stationarity
#it doesn't get better when I remove dyads with 0 aid total

#these show that the data is stationary

adf_sample <- replicate(100, purtest(adf_mat[ ,sample(2:2160, 50)],
                            pmax = 5, 
                            exo = "trend", 
                            test = "madwu",
                            lags = "AIC")[[1]][6])


```


Manual ADF
```{r}


adf_m <- lm(democracy_aid_com_corr_df ~ 
                 democracy_aid_com_corr_lag +
                 democracy_aid_com_corr_df_lag +
                 ccode1 + 
                 ccode2 + 
                 as.factor(year), 
               data = mergepdata4,
               na.action = na.omit)

summary(adf_m)

adf_m_hc3 <- coeftest(adf_m, cluster.vcov(adf_m,
                                         cluster = cbind(mergepdata4$ccode1, 
                                                         mergepdata4$ccode2), 
                                         leverage = 3, 
                                         df_correction = T))[-(24:nrow(summary(adf_m)$coefficients)),]; adf_m_hc3

#these results also show that the data is stationary (which is good)


```

***



Testing for autocorrelation - maybe not include this

```{r}

# ## Dec 20 - for panel data, autocorrelation can be handled by robust standard errors
# 
# # https://www.statalist.org/forums/forum/general-stata-discussion/general/1390195-heteroskedasticity-serial-correlation-in-panel-data-with-fixed-effects
# 
# 
# #####
# table(model2aa$residuals == resid(model2aa))
# 
# mergepdata$error <- resid(model2aa)
# mergepdata$l.error <- ave(mergepdata$error,
#                          mergepdata$ccode1,
#                          mergepdata$ccode2,
#                          FUN=function(x) lagpad(x, 1))
# 
# ### Auxiliary regression
# model2_aux <- lm(error ~ l.error,
#                  data = mergepdata,
#                  na.action = na.exclude)
# 
# m2auxhc3 <- coeftest(model2_aux, cluster.vcov(model2_aux,
#                                            cluster = cbind(mergepdata$ccode1, mergepdata$ccode2), 
#                                            leverage = 3, 
#                                            df_correction = T))
# #the lagged error here is significant, so there IS autocorrelation

```  


