---
title: "panel_data_building"
author: "SaewonPark"
date: "April 23, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

pacman::p_load(tidyverse,
               countrycode,
               haven,
               readxl,
               plm,
               RcppRoll) 


lag <- dplyr::lag
lead <- dplyr::lead

```

# Building a Panel dataset 

#### Info

My base dataset was the CRS (aid information from OECD) dataset
*year range of CRS dataset: 1973-2016

However, we limit the time range to 1990-2014 (due the timeframe of the controls dataset)



###CRS DATASET

```{r}
######Import data

##2016
crs2016 <- read.csv("CRS 2016 data.txt", row.names=NULL, sep="|")

##2015
crs2015 <- read.csv("CRS 2015 data.txt", row.names=NULL, sep="|")

##2014
crs2014 <- read.csv("CRS 2014 data.txt", row.names=NULL, sep="|")

##2013
crs2013 <- read.csv("CRS 2013 data.txt", row.names=NULL, sep="|")

##2012
crs2012 <- read.csv("CRS 2012 data.txt", row.names=NULL, sep="|")

##2011
crs2011 <- read.csv("CRS 2011 data.txt", row.names=NULL, sep="|")

##2010
crs2010 <- read.csv("CRS 2010 data.txt", row.names=NULL, sep="|",  quote = "")

##2009
crs2009 <- read.csv("CRS 2009 data.txt", row.names=NULL, sep="|",  quote = "")

##2008
crs2008 <- read.csv("CRS 2008 data.txt", row.names=NULL, sep="|",  quote = "")

##2007
crs2007 <- read.csv("CRS 2007 data.txt", row.names=NULL, sep="|",  quote = "")

##2006
crs2006 <- read.csv("CRS 2006 data.txt", row.names=NULL, sep="|",  quote = "")

##2004-5
crs2004_05 <- read.csv("CRS 2004-05 data.txt", row.names=NULL, sep="|",  quote = "")

##2002-3
crs2002_03 <- read.csv("CRS 2002-03 data.txt", row.names=NULL, sep="|", quote = "")

##2000-1
crs2000_01 <- read.csv("CRS 2000-01 data.txt", row.names=NULL, sep="|",  quote = "")

##1995-1999
crs1999_95 <- read.csv("CRS 1995-99 data.txt", row.names=NULL, sep="|",  quote = "")

##1973-1994
crs1973_94 <- read.csv("CRS 1973-94 data.txt", row.names=NULL, sep="|",  quote = "")
```



Prior to the Dec 2018 edits, a lot of functions that could have been piped had not been piped

```{r}

##########Bind all the data together##########

#the column "interest1" causes problems and we don't need it so I just drop it
#oct 6: also dropped other variables to reduce size
listcrs <- mget(ls(pattern='^crs\\d+'))
listcrs <- lapply(listcrs, function(x) x[!grepl(paste(c("*interest1", 
                                                           "*biodiversity",
                                                           "*climateMitigation",
                                                           "*desertification",
                                                           "*climateAdaptation",
                                                           "*Gender",
                                                           "*Environment"), collapse = "|"), names(x))])
                  
list2env(listcrs, envir = .GlobalEnv)
rm(listcrs)

gc()

rm(list=ls())

#merging data

#because r kept crashing, I didn't bind everything at once
crspanel <- bind_rows(crs1973_94,
                      crs1999_95,
                      crs2000_01,
                      crs2002_03)
write.csv(crspanel, "./crs/crs_73-03.csv", row.names = F)
rm(list=ls())

crspanel <- bind_rows(crs2004_05,
                      crs2006,
                      crs2007)
write.csv(crspanel, "./crs/crs_04-07.csv", row.names = F)
rm(list=ls())

crspanel <- bind_rows(crs2008,
                      crs2009,
                      crs2010)
write.csv(crspanel, "./crs/crs_08-10.csv", row.names = F)
rm(list=ls())

crspanel <- bind_rows(crs2011,
                      crs2012,
                      crs2013)
write.csv(crspanel, "./crs/crs_11-13.csv", row.names = F)
rm(list=ls())

crspanel <- bind_rows(crs2014,
                      crs2015,
                      crs2016)

write.csv(crspanel, "./crs/crs_14-16.csv", row.names = F)
rm(list=ls())

# https://stackoverflow.com/questions/27399814/how-to-remove-certain-columns-in-multiple-data-frames-in-r?utm_medium=organic&utm_source=google_rich_qa&utm_campaign=google_rich_qa
# https://stackoverflow.com/questions/12664430/delete-a-column-in-a-data-frame-within-a-list?utm_medium=organic&utm_source=google_rich_qa&utm_campaign=google_rich_qa

crs_73_03 <- read.csv("./crs/crs_73-03.csv", stringsAsFactors = F)
crs_04_07 <- read.csv("./crs/crs_04-07.csv", stringsAsFactors = F)

crspanel <- bind_rows(crs_73_03,
                      crs_04_07)
write.csv(crspanel, "./crs/crs_73-07.csv", row.names = F)
rm(list=ls())

crs_08_10 <- read.csv("./crs/crs_08-10.csv", stringsAsFactors = F)
crs_11_13 <- read.csv("./crs/crs_11-13.csv", stringsAsFactors = F)
crs_14_16 <- read.csv("./crs/crs_14-16.csv", stringsAsFactors = F)

#fixing column names for pre-2010 data
names(crs_08_10) <- gsub("X.", "",  names(crs_08_10)) %>%
  gsub('.{1}$', '', .)

crspanel <- bind_rows(crs_08_10,
                      crs_11_13,
                      crs_14_16) %>%
  select(-LongDescription)

write.csv(crspanel, "./crs/crs_08-16.csv", row.names = F)
rm(list=ls())


### binding everything together

crs_73_07 <- read.csv("./crs/crs_73-07.csv", stringsAsFactors = F)
names(crs_73_07) <- gsub("X.", "",  names(crs_73_07)) %>%
  gsub('.{1}$', '', .)
crs_73_07 <- crs_73_07 %>% 
  select(-LongDescription)

crs_08_16 <- read.csv("./crs/crs_08-16.csv", stringsAsFactors = F)

crspanel <- bind_rows(crs_73_07,
                      crs_08_16)
write.csv(crspanel, "./crs/crspanel.csv", row.names = F)
rm(crs_73_07, crs_08_16)

## we also need to remove the column LongDescription (which contains a paragraph each and is taking up too much space)


######## Change country name to cowcode ########
#we do this because OECD has their own country code 

##donors##

countries_cow_crsdonor <- unique(countrycode(crspanel$DonorName, "country.name", "cown", warn = TRUE ))
countries_cow_crsdonor <- countries_cow_crsdonor[!countries_cow_crsdonor == ""]
countries_cow_crsdonor <- countries_cow_crsdonor[!is.na(countries_cow_crsdonor)]

crspanel$ccodedonor <- countrycode(crspanel$DonorName, "country.name", "cown", warn = TRUE )

##recipients##
countries_cow_crsrecipient <- unique(countrycode(crspanel$RecipientName, "country.name", "cown", warn = TRUE ))
countries_cow_crsrecipient <- countries_cow_crsrecipient[!countries_cow_crsrecipient == ""]
countries_cow_crsrecipient <- countries_cow_crsrecipient[!is.na(countries_cow_crsrecipient)]

crspanel$ccoderecipient <- countrycode(crspanel$RecipientName, "country.name", "cown", warn = TRUE )
#we need to code Serbia separately, because cow code uses the same code for the former Yugoslavia
crspanel$ccoderecipient[crspanel$RecipientName == "Serbia"] <- 345 


##let's make sure that only the non-countries or very small countries (islands) will be removed
#donors that will be removed
as.vector(unique(filter(crspanel, is.na(ccodedonor))$DonorName))

#recipients that will be removed
as.vector(unique(filter(crspanel, is.na(ccoderecipient))$RecipientName))
#looks good



####### Clean the CRS data ########
##Info about CRS data: http://www.oecd.org/dac/stats/crsguide.htm

crspanel <- read.csv("./crs/crspanel.csv", stringsAsFactors = F)

#deflators in 2016 constant USD
deflators <- read_excel("../oecd/Deflators-base-2016.xls", skip = 2) %>%
  rename(countryname = X__1) %>%
  select(-X__2) %>%
  filter(!is.na(`1960`) & 
           !(countryname %in% c("TOTAL DAC", "EU Institutions"))) %>%
  gather(year, deflator, -countryname) %>%
  mutate(ccode1 = countrycode(countryname, "country.name", "cown"),
         deflator = deflator/100,
         year = as.numeric(year))

  



#let's get an idea of the other aid
aid_type_count <- crspanel %>% 
  select(usd_disbursement,
         SectorCode,
         SectorName) %>%
  group_by(SectorName, SectorCode) %>%
  summarise(count = n(),
            usd_disbursement = sum(usd_disbursement, na.rm = T)) %>%
  filter(SectorCode != 998 & #998 is for unspecified aid
           SectorName != "") %>% #one field was left blank (very small, 10 count)
  mutate(aidtype = case_when(SectorCode %in% c(100:140, 152, 160) ~ "social_aid",
                             SectorCode %in% c(150,151) ~ "democracy_aid",
                             SectorCode %in% c(200:250) ~ "econ_capacity_aid",
                             SectorCode %in% c(300:332) ~ "production_aid",
                             SectorCode %in% c(500:600) ~ "program_aid",
                             SectorCode %in% c(720:740) ~ "disaster_aid",
                             TRUE ~ "other_aid")) %>%
  ungroup()
#checked that all channels(sub-sectors) belong to some sector

#quick pie chart
ggplot(aid_type_count, aes(x = "", y = usd_disbursement, fill = aidtype)) +
  geom_bar(stat = "identity") +
  coord_polar("y", start=0)


#### begin Dec 16 to crs data
#we need all sectors of aid

crsall <- left_join(crspanel,
                    select(aid_type_count,
                           SectorCode,
                           aidtype),
                    by = "SectorCode") %>%
  filter(!is.na(aidtype) &
           SectorCode != 998 & #998 is for unspecified aid
           SectorName != "" & #one field name was left blank (very small, 10 count))
           !is.na(ccodedonor) & 
           !is.na(ccoderecipient) & #Filter out all the non-countries
           !(is.na(usd_disbursement) & is.na(usd_commitment))) %>% 
  select(Year,
         ccodedonor,
         ccoderecipient,
         aidtype,
         usd_disbursement,
         usd_commitment) %>%
  rename(year = Year,
         ccode1 = ccodedonor,
         ccode2 = ccoderecipient,
         aid_disburse = usd_disbursement,
         aid_commit = usd_commitment) %>%
  left_join(select(deflators,
                   year, 
                   ccode1,
                   deflator),
            by = c("year", "ccode1")) %>% #deflating aid according to CRS deflators for donor-year
  mutate(aid_disburse = aid_disburse/deflator,
         aid_commit = aid_commit/deflator)


  
#divide into disbursement and commitments, spread into sectors and then merge again
crsall_dis <- crsall %>%
  select(year,
         ccode1,
         ccode2,
         aid_disburse,
         aidtype) %>%
  filter(!is.na(aid_disburse)) %>%
  group_by(ccode1, ccode2, year, aidtype) %>%
  summarise(aid_disburse = sum(aid_disburse, na.rm = T)) %>%
  spread(aidtype, aid_disburse, fill = 0) %>% #because 0 aid was not recorded (all NAs were already removed)
  rename_at(vars(-ccode1, -ccode2, -year), funs(paste0(.,"_disb"))) %>%
  ungroup() %>%
  group_by(ccode1, ccode2) %>%
  complete(year = 1973:2016, fill = list(democracy_aid_disb = 0,
                                         disaster_aid_disb = 0,
                                         econ_capacity_aid_disb = 0,
                                         other_aid_disb = 0,
                                         production_aid_disb = 0,
                                         program_aid_disb = 0,
                                         social_aid_disb = 0))


crsall_com <- crsall %>%
  select(year,
         ccode1,
         ccode2,
         aid_commit,
         aidtype) %>%
  filter(!is.na(aid_commit)) %>%
  group_by(ccode1, ccode2, year, aidtype) %>%
  summarise(aid_commit = sum(aid_commit, na.rm = T)) %>%
  spread(aidtype, aid_commit, fill = 0) %>% #because 0 aid was not recorded (all NAs were already removed)
  rename_at(vars(-ccode1, -ccode2, -year), funs(paste0(.,"_com"))) %>%
  ungroup() %>%
  group_by(ccode1, ccode2) %>%
  complete(year = 1973:2016, fill = list(democracy_aid_com = 0,
                                         disaster_aid_com = 0,
                                         econ_capacity_aid_com = 0,
                                         other_aid_com = 0,
                                         production_aid_com = 0,
                                         program_aid_com = 0,
                                         social_aid_com = 0))
  

#now merge the spread data for both commitments and disbursements
crsall_sp <- full_join(crsall_com,
                       crsall_dis,
                       by = c("ccode1",
                              "ccode2",
                              "year")) %>%
  filter(year > 1985) %>%
  replace_na(list(democracy_aid_com = 0,
             democracy_aid_disb = 0))
  


### end dec 16 edit


# I will limit my dataset to a timeframe of 1990-2014
#crscollapsed <- filter(crscollapsed, year > 1989 & year < 2015)

#looks good
```



### IDEAL POINTS DATASET

* I saved the rmd as a new dataset on dec 16

```{r}
#1946-2014: dyad data (has ideal points and affinity data (s2un, s3un)- refer to code book)
  #absidealdiff in dyad data contains the absolute distance between country 1 and country 2 posterior mean ideal point estimates.
  #uses COW codes

#####Import data
dyaddata <- read.table(
  "Dyadicdata.tab",
  sep="\t", header=TRUE)

##### Clean data #####
#absidealdiff is in the opposite direction of s3un
#greater absidealdiff and lesser s3un means greater difference

## Dec 16
dyadclean <- select(dyaddata, ccode1, ccode2, year, absidealdiff) %>%
  rename(ipdiff = absidealdiff) %>% 
  group_by(ccode1, ccode2) %>%
  arrange(ccode1, ccode2, year) %>%
  mutate(ipfd = ipdiff - lag(ipdiff, order_by = year), #first diff
         ipdiff_1 = lag(ipdiff, order_by = year), # Ideal points diff- 1 yr, 3 yr, 5 yr (savun uses 5)
         ipdiff_3 = lag(ipdiff, order_by = year, n = 3),
         ipdiff_5 = lag(ipdiff, order_by = year, n = 5), # Similarly, ideal points FD- 1 yr, 3 yr, 5 yr
         ipfd_1 = lag(ipfd, order_by = year),
         ipfd_3 = lag(ipfd, order_by = year, n = 3),
         ipfd_5 = lag(ipfd, order_by = year, n = 5)) 
  

```



##### CRS and I.P. MERGE 

```{r}

mergedata1 <- left_join(crsall_sp,
                       dyadclean, 
                       by = c("ccode1", "ccode2", "year"))

```



### Controls 

```{r}
##### Adding Controls
  # these ones are already lagged one year

controls <- read_dta("extended_data_0401.dta") %>%
  mutate(year = as.integer(year),
         cowcode = as.integer(cowcode))

#already lagged one year
#in country-year



###### Adding control variables to the dataset ######

###Recipient
#controls$conflict_prioryear
#controls$polity2_dem_ord #this is the change in democratization
#controls$democracy #this is the democracy score
#controls$pop #this is the population that has not been logged #I use it to scale the aid flow

###Both
#controls$growth_real_gdppc
#controls$real_gdppc
#controls$population

###Donor
#controls$consumer_inflation
#controls$women_parliament

#remember that pop and aidflows are both in millions

mergedata2 <- left_join(mergedata1,
                       select(controls,
                              cowcode,
                              year,
                              conflict_prioryear,
                              pop, #this is just used to correct the aid flow - this is in millions
                              polity2_dem_ord,
                              polity2,
                              growth_real_gdppc,
                              real_gdppc,
                              population), #"population"" is logged pop in thousands
                       by = c("ccode2" = "cowcode", 
                              "year"="year")) %>%
  rename(r_democracy = polity2,
         r_conflict_prioryear = conflict_prioryear,
         r_pop_notlogged = pop,
         r_politychange = polity2_dem_ord,
         r_gdp_growth = growth_real_gdppc,
         r_gdp = real_gdppc,
         r_population = population) %>%
  mutate(r_democratization = as.numeric(r_politychange > 2),
         r_gdplog = log(r_gdp),
         r_democracy_sq = (r_democracy)^2) %>%
  left_join(select(controls,
                   cowcode,
                   year,
                   pop,
                   growth_real_gdppc,
                   real_gdppc,
                   population,
                   consumer_inflation,
                   women_parliament),
            by = c("ccode1" = "cowcode", 
                   "year"="year")) %>%
  rename(d_gdp_growth = growth_real_gdppc,
         d_pop_notlogged = pop,
         d_gdp = real_gdppc,
         d_population = population,
         d_consumer_inflation = consumer_inflation,
         d_women_parliament = women_parliament) %>%
  mutate(d_gdplog = log(d_gdp)) %>%
  mutate_at(vars(contains("aid")), 
            .funs = funs(corr = log(((./r_pop_notlogged)*1000)+1))) %>%
  group_by(ccode2, ccode1) %>%
  mutate_at(vars(ends_with("corr")),
            .funs = funs(`1` = lag(., order_by = year))) %>%
  mutate_at(vars(ends_with("corr")),
            .funs = funs(`3` = lag(., order_by = year, n = 3))) %>%
  mutate_at(vars(ends_with("corr")),
            .funs = funs(`5` = lag(., order_by = year, n = 5))) %>%
  mutate_at(vars(starts_with("r_")),
            .funs = funs(`1` = lag(., order_by = year))) %>%
  mutate_at(vars(starts_with("d_")),
            .funs = funs(`1` = lag(., order_by = year)))

#NA created for aid corrected for population because of missing population data   
  
#sanity check - only very few are nans
#sapply(mergedata, function(x) sum(is.nan(x)))
#head(select(mergedata, democracy_aid_com_corr, democracy_aid_com_corr_l1)[100:106,])

#for 666, lag democratization by 3 - should I still do this?

write.csv(mergedata2, "mergedata_dec16.csv", row.names = F)

```


```{r}
###CEPII data for distance (no year) 
cepii <- read_dta("dist_cepii.dta") %>%
  mutate(ccode1 = countrycode(iso_o, "iso3c", "cown"),
         ccode2 = countrycode(iso_d, "iso3c", "cown"),
         distlog = log(dist),
         d_region = countrycode(ccode1, "cown", "region"),
         r_region = countrycode(ccode2, "cown", "region"),
         samereg = as.numeric(d_region == r_region)) 


#only the small countries (islands) are missing
#dist is the simple distance between the biggest cities
#colony is whether the pair of countries were ever in a colonial relationship
#distw and distwces are basically the same, but distwces is the inverse to distw
  #both are weighted distance
  #distwces shrinks as distance grows and distw is the opposite



```


### DESTA data- PTA 

```{r}
#create column for number of treaties to be filled with 0s too after merge (?- april 19)
#for some reason I only looked at PTAs prior to 2005 - idk why I did that, so I removed this restriction

DESTA <- read.csv(file="DESTA_treaties_dyadic.csv", stringsAsFactors = F)

desta1 <- select(DESTA, 
                 iso1, 
                 iso2, 
                 base_treaty,
                 year) %>%
  mutate(ccode1 = countrycode(iso1, "iso3n", "cown"),
         ccode2 = countrycode(iso2, "iso3n", "cown")) %>%
  filter(!(is.na(ccode1)) & !is.na(ccode2)) %>%
  distinct() %>%
  group_by(ccode1, ccode2, year) %>%
  summarise(ptacount = n()) 


desta_both <- rbind(desta1 %>%
                      ungroup(),
                    desta1 %>%
                      rename(c1 = ccode1,
                             c2 = ccode2) %>%
                      mutate(ccode1 = c2,
                             ccode2 = c1) %>%
                      ungroup() %>%
                      select(-c1, -c2)) %>%
  group_by(ccode1, ccode2, year) %>%
  summarise(ptacount = n()) %>%
  ungroup()

  

#ptacount equals 1 for both direction of the order of ccode1/2
  #in desta, each agreement is only counted once (if there is US-CAN, then there is no CAN-US for a particular agreement)

#because there are missing years, the cumulative value and lagged values need to be calculated after merge
```




### Unemployment Data 

```{r}
unemployment_worldbank <- read.csv(file="Unemployment_WB.csv", skip = 4, header=TRUE, sep=",") %>%
  select(-X)


#remove the "X" in front of all the column names for the years

unemployment <- unemployment_worldbank %>%
  select(-Indicator.Code,
         -Indicator.Name) %>% #we only picked one indicator so this is all the same
  rename_at(vars(starts_with("X")),
                 funs(substr(.,2, 5))) %>%
  gather(year, unemployment, -c(Country.Name, Country.Code)) %>%
  group_by(Country.Code) %>%
  mutate(year = as.integer(year),
         ccode = countrycode(Country.Code, "iso3c", "cown"),
         unemployment_1 = lag(unemployment, order_by = year)) %>% #lag 1 year
  ungroup()


#looks good
```



### Immigration Data 

```{r}
immigration_oecd <- read.csv("Migration_1975-2015.csv", stringsAsFactors = F)

#not scaled (count at single person level, not in millions, etc.)
#all host countries are OECD members

#filter only "inflow of foreign population by nationality" (to an OECD member), or VAR = B11
  #foreign population means people with another nationality

#DEC 12 : also filter B16 (Acquisition of nationality by country of former nationality)
  #this is naturalization data 

immigration <- filter(immigration_oecd, 
                      VAR %in% c("B11", "B16")) %>%
  select(`ï..CO2`,
         COU,
         Year,
         Value,
         VAR) %>% 
  spread(VAR, Value) %>%
  rename(country1 = COU, #donor (inflow)
         country2 = `ï..CO2`, #recipient (outflow)
         year = Year,
         immigration = B11,
         naturalization = B16) %>%
  mutate(ccode1 = countrycode(country1, "iso3c", "cown"),
         ccode2 = countrycode(country2, "iso3c", "cown")) %>%
  filter(!is.na(ccode1) & !is.na(ccode2)) %>%
  group_by(country1,
           country2) %>%
  mutate(immigration_1 = lag(immigration, order_by = year),
         immigration_3 = lag(immigration, 3, order_by = year),
         immigration_5 = lag(immigration, 5, order_by = year),
         immigration_7 = lag(immigration, 7, order_by = year),
         immigration_10 = lag(immigration, 10, order_by = year),
         naturalization_1 = lag(naturalization, order_by = year),
         naturalization_3 = lag(naturalization, 3, order_by = year),
         log_immigration = log(immigration + 1),
         log_immigration_1 = log(immigration_1 + 1),
         log_immigration_3 = log(immigration_3 + 1),
         log_immigration_5 = log(immigration_5 + 1),
         log_immigration_7 = log(immigration_7 + 1),
         log_immigration_10 = log(immigration_10 + 1),
         log_naturalization = log(naturalization + 1),
         log_naturalization_1 = log(naturalization_1 + 1),
         log_naturalization_3 = log(naturalization_3 + 1)) %>%
  ungroup()


#the country1 column includes TOT which means "total"- we should remove this


#I lag this 1, 3, and 5 years because there is an argument in Menard that immigration can lead to unemployment 
# which can lead to aid (I don't think this is directly related to lagging, but it will be good to check in case)


#Dec 12 - I also did a 7 year lag (used by Jones-Correa) and a 10 year lag to proxy for immigrants who have stayed for a while and are more likely to have been naturalized. 


#Dec 29 - need to check that all years are present and the lags actually work- it does, the lags are ok
 # - should use log_naturalization_lag


#summary(immigration$immigration)
#looks good


# corrplot(cor(select(immigration,
#                     immigration,
#                     naturalization), use = "complete.obs"),
#          method = "number")

#here the correlation between immigration and naturalization is not too correlated

#sanity check
#head(immigration, n = 10)

```



### Trade Data from IMF

```{r}
#I use exports like DESTA did
  #use export from oecd to recipient, and then recipient to oecd
#export in dollars

trade_imf <- read.csv(file="DOT_04-19-2018 19-48-46-85_panel.csv",
                      stringsAsFactors = F)

#select just the columns I need
trade <- select(trade_imf,
                Country.Code, 
                Counterpart.Country.Code, 
                Time.Period,
                Goods..Value.of.Exports..Free.on.board..FOB...US.Dollars..TXG_FOB_USD.) %>%
  rename(countryfrom = Country.Code,
         countryto = Counterpart.Country.Code,
         year = Time.Period,
         export = Goods..Value.of.Exports..Free.on.board..FOB...US.Dollars..TXG_FOB_USD.) %>%
  mutate(ccodefrom = countrycode(countryfrom, "imf", "cown"),
         ccodeto = countrycode(countryto, "imf", "cown")) %>% #change from IMF code
  filter(!is.na(ccodeto) & !is.na(ccodefrom))



```


### US military aid from foreign aid greenbook

```{r}

#in 2016 constant dollars, obligations not disbursement

us_military_aid <- read_excel("us_foreignaid_greenbook.xlsx", skip = 6) %>%
  rename(year = `Fiscal Year`,
         country_name = Country,
         category = `Assistance Category`,
         military_aid = `Obligations (Constant Dollars)`) %>%
  filter(category == "Military") %>%
  select(year,
         country_name,
         military_aid) %>%
  mutate(ccode2 = countrycode(country_name, "country.name", "cown")) %>%
  filter(!is.na(ccode2)) %>%
  group_by(year,
           ccode2) %>%
  summarise(military_aid = sum(military_aid)) %>%
  ungroup() %>%
  group_by(ccode2) %>%
  complete(year = 1947:2016) %>%
  mutate(military_aid = ifelse(is.na(military_aid), 0, military_aid))



```




Merging part 3- final

```{r}


mergedata3 <- left_join(mergedata2,
                        select(cepii,
                               contig,
                               comlang_off,
                               comlang_ethno,
                               colony,
                               dist,
                               distw,
                               distwces,
                               ccode1,
                               ccode2,
                               distlog, #this is the log of simple distance
                               d_region,
                               r_region,
                               samereg),
                        by = c("ccode1",
                               "ccode2")) %>%
  left_join(desta_both,
            by = c("ccode1",
                   "ccode2",
                   "year")) %>%
  group_by(ccode1, ccode2) %>%
  arrange(ccode1, ccode2, year) %>%
  mutate(ptacount = ifelse(is.na(ptacount), 0, ptacount), #added up duplicates already
         ptasum = cumsum(ptacount), # Cumulative count of treaties - will need to autocorrelation
         ptasum_1 = lag(ptasum, order_by = year),
         ptacount_1 = lag(ptacount, order_by = year)) %>%
  left_join(select(unemployment, #unemployment columns for recipient
                   ccode,
                   year,
                   unemployment,
                   unemployment_1),
            by = c("ccode2" = "ccode",
                   "year")) %>%
  rename(r_unemployment = unemployment,
         r_unemployment_1 = unemployment_1) %>%
  left_join(select(unemployment, #unemployment columns for donors
                   ccode,
                   year,
                   unemployment,
                   unemployment_1),
            by = c("ccode1" = "ccode",
                   "year")) %>%
  rename(d_unemployment = unemployment,
         d_unemployment_1 = unemployment_1) %>%
  mutate(unempdiff = d_unemployment - r_unemployment, #also a column for unemployment differences
         unempdiff_1 = lag(unempdiff, order_by = year)) %>% #negative value: donor has less unemployment
  left_join(select(immigration,
                   -country1,
                   -country2),
            by = c("ccode1",
                   "ccode2",
                   "year")) %>% #adding a 5-year rolling sum of naturalization
  mutate(naturalization_roll = roll_sum(naturalization, 5, fill = NA, align = "right"),
         log_naturalization_roll = log(roll_sum(naturalization, 5, fill = NA, align = "right")+1),
         log_naturalization_roll_1 = lag(log_naturalization_roll, order_by = year),
         immigration_roll = roll_sum(immigration, 5, fill = NA, align = "right"),
         log_immigration_roll = log(roll_sum(immigration, 5, fill = NA, align = "right")+1),
         log_immigration_roll_1 = lag(log_immigration_roll, order_by = year)) %>%
  left_join(select(trade, #need to merge separately to get d to r and r to d exports
                   ccodeto,
                   ccodefrom,
                   year,
                   export),
            by = c("ccode1" = "ccodefrom",
                   "ccode2" = "ccodeto",
                   "year")) %>%
  rename(dtor_export = export) %>%
  left_join(select(trade, #need to merge separately to get d to r and r to d exports
                   ccodeto,
                   ccodefrom,
                   year,
                   export),
            by = c("ccode2" = "ccodefrom",
                   "ccode1" = "ccodeto",
                   "year")) %>%
  rename(rtod_export = export) %>%
  mutate(log_rtod_export = log(rtod_export+1),
         log_dtor_export = log(dtor_export+1),
         rtod_export_1 = lag(rtod_export, order_by = year),
         dtor_export_1 = lag(dtor_export, order_by = year),
         log_rtod_export_1 = lag(log_rtod_export, order_by = year),
         log_dtor_export_1 = lag(log_dtor_export, order_by = year)) %>%
  left_join(us_military_aid,
            by = c("ccode2",
                   "year")) %>%
  mutate(military_aid_corr = log((military_aid/r_pop_notlogged*1000)+1),
         military_aid_corr_1 = lag(military_aid_corr, order_by = year),
         military_aid_corr_3 = lag(military_aid_corr, 3, order_by = year),
         military_aid_corr_5 = lag(military_aid_corr, 5, order_by = year),
         dyad_id = paste(ccode1, ccode2, sep = "_")) #creating a dyad-id for augmented dickey fuller tests


######################## DONE with merging ######################
#################################################################


###to do TWC with HC3 errors, turn data into pdata.frame and use plm
mergepdata <- pdata.frame(mergedata3, index = c("ccode1", "ccode2"))

#there are warning because of duplicate dyad (since I have years now)
#let's proceed for now and see what happens

#sanity check
head(mergedata3[200:210,c(1,2,3,133,134,142,148,151)], n = 10)


```





### Corrections, other cleaning tasks


```{r}
########## I need to remove all missing aidflow corrected value#######
  #since we can't have missing Y

# mergedata <- filter(mergedata, 
#                      !is.na(aidflow_corrected) & !is.nan(aidflow_corrected) & !is.infinite(aidflow_corrected))
  # I would not do this because we still need to lag data


######### I also need to double lag all variables and get the FD for ECM  ########
  #ipfd would be dropped and added as fd
  

#for testing for stationarity and first difference model

mergedata4 <- mergedata3 %>%
  group_by(ccode1, ccode2) %>%
  mutate(naturalization_prop_log = log(naturalization/(d_pop_notlogged)+1), #proportional to donor pop in mil
         immigration_prop_log = log(immigration/(d_pop_notlogged)+1),
         naturalization_roll_prop_log = log(naturalization_roll/(d_pop_notlogged)+1),
         immigration_roll_prop_log = log(immigration_roll/(d_pop_notlogged)+1)) %>%
  mutate_all(.funs = funs(lag = lag(., order_by = year))) %>%
  select(-dyad_id_lag) %>%
  mutate(democracy_aid_com_corr_df = democracy_aid_com_corr - democracy_aid_com_corr_lag,
         democracy_aid_disb_corr_df = democracy_aid_disb_corr - democracy_aid_disb_corr_lag,
         democracy_aid_com_corr_df_lag = lag(democracy_aid_com_corr_df, order_by = year),
         democracy_aid_disb_corr_df_lag = lag(democracy_aid_disb_corr_df, order_by = year),
         disaster_aid_com_corr_df = disaster_aid_com_corr - disaster_aid_com_corr_lag,
         econ_capacity_aid_com_corr_df = econ_capacity_aid_com_corr - econ_capacity_aid_com_corr_lag,
         other_aid_com_corr_df = other_aid_com_corr - other_aid_com_corr_lag,
         production_aid_com_corr_df = production_aid_com_corr - production_aid_com_corr_lag,
         program_aid_com_corr_df = program_aid_com_corr - program_aid_com_corr_lag,
         social_aid_com_corr_df = social_aid_com_corr - social_aid_com_corr_lag,
         disaster_aid_disb_corr_df = disaster_aid_disb_corr - disaster_aid_disb_corr_lag,
         econ_capacity_aid_disb_corr_df =econ_capacity_aid_disb_corr - econ_capacity_aid_disb_corr_lag,
         other_aid_disb_corr_df = other_aid_disb_corr - other_aid_disb_corr_lag,
         production_aid_disb_corr_df = production_aid_disb_corr - production_aid_disb_corr_lag,
         program_aid_disb_corr_df = program_aid_disb_corr - program_aid_disb_corr_lag,
         social_aid_disb_corr_df = social_aid_disb_corr - social_aid_disb_corr_lag,
         r_conflict_prioryear_df = r_conflict_prioryear - r_conflict_prioryear_lag, #I skip ipdiff bc it is already fd
         r_politychange_df = r_politychange - r_politychange_lag,
         r_democracy_df = r_democracy - r_democracy_lag,
         r_gdp_growth_df = r_gdp_growth - r_gdp_growth_lag,
         r_gdp_df = r_gdp - r_gdp_lag,
         r_population_df = r_population - r_population_lag,
         r_democratization_df = r_democratization - r_democratization_lag,
         r_gdplog_df = r_gdplog - r_gdplog_lag,
         r_democracy_sq_df = r_democracy_sq - r_democracy_sq_lag,
         d_gdp_growth_df = d_gdp_growth - d_gdp_growth_lag,
         d_gdp_df = d_gdp - d_gdp_lag,
         d_population_df = d_population - d_population_lag,
         #d_consumer_inflation_df = d_consumer_inflation - d_consumer_inflation_lag,  # this a chr vector
         d_women_parliament_df = d_women_parliament - d_women_parliament_lag,
         d_gdplog_df = d_gdplog - d_gdplog_lag,
         ptacount_df = ptacount - ptacount_lag,
         ptasum__df = ptasum - ptasum_lag,
         unempdiff_df = unempdiff - unempdiff_lag,
         log_immigration_df = log_immigration - log_immigration_lag,
         log_immigration_5_df = log_immigration_5 - log_immigration_5_lag,
         log_naturalization_df = log_naturalization - log_naturalization_lag,
         log_naturalization_1_df = log_naturalization_1 - log_naturalization_1_lag,
         log_naturalization_3_df = log_naturalization_3 - log_naturalization_3_lag,
         log_naturalization_roll_df = log_naturalization_roll - log_naturalization_roll_lag,
         log_immigration_roll_lag = log_immigration_roll - log_immigration_roll_lag,
         log_rtod_export_df = log_rtod_export - log_rtod_export_lag,
         log_dtor_export_df = log_dtor_export - log_dtor_export_lag,
         military_aid_df = military_aid - military_aid_lag,
         military_aid_corr_df = military_aid_corr - military_aid_corr_lag,
         other_economic_aid = log(((disaster_aid_disb +
                                    econ_capacity_aid_disb +
                                    other_aid_disb + 
                                    production_aid_disb +
                                    program_aid_disb +
                                    social_aid_disb)/1000)+1),
         other_economic_aid_lag = lag(other_economic_aid, order_by = year),
         other_economic_aid_df = other_economic_aid - other_economic_aid_lag,
         naturalization_prop_log_df = naturalization_prop_log - naturalization_prop_log_lag,
         immigration_prop_log_df = immigration_prop_log - immigration_prop_log_lag,
         naturalization_roll_prop_log_df = naturalization_roll_prop_log - naturalization_roll_prop_log_lag,
         immigration_roll_prop_log_df = immigration_roll_prop_log - immigration_roll_prop_log_lag) %>%
  mutate_at(vars(ends_with("_df")),
            .funs = funs(lag = lag(., order_by = year)))

#str(mergedata4[ , 250:338])
 #I should lag all the dfs (DONE)

```







```{r}

# ###no iraq data
# #remove the US and Iraq (or just Iraq)
# mergepdata_noirq <- filter(mergepdata, ccode2 != 625 )
# #countrycode("Iraq", "country.name", "cown")
# 
# #looks good
# 
# 
# #let's export this as *.dta for later use
#   # we need to take care of factors before converting to dta
# 
# mergepdata$iso_o <- as.character(mergepdata$iso_o)
# mergepdata$iso_d <- as.character(mergepdata$iso_d)
# 
# mergepdata$region1 <- as.character(mergepdata$region1)
# mergepdata$region2 <- as.character(mergepdata$region2)
# 
# mergepdata$continent1 <- as.character(mergepdata$continent1)
# mergepdata$continent2 <- as.character(mergepdata$continent2)
# 
# class(mergepdata$region1)
# 
# table(sapply(mergepdata, is.factor))
# 
# write.dta(mergepdata, "C:/Users/Spark/Desktop/Winter_2018/ORGB_708/research/researchproj/708data_0421.dta", 
#           convert.factors = "numeric")
# 
# 
# #added oct 6 2018
# aa <- read_dta("708data_0421.dta")
# 
# write.csv(aa, "UNGA_fulldata.csv", row.names = F)


```






